---
title-block-banner: true
title: |
  Data-Intensive Social Science Center (DISSC) Curated Social Science Data
description: |
  </br>
  Founded in 2022, Yale's <a href = "https://dissc.yale.edu/">Data-Intensive Social Science Center (DISSC)</a> serves as a central hub supporting data-intensive social science research across the university, providing access to computational resources, data science tools, and statistical methods. In partnership with the Yale Library, DISSC maintains a curated listing of licensed and publicly available datasets, including census microdata, consumer spending, and health and pharmacy data, accessible through their <a href = "https://dissc.yale.edu/data-resources">Discover and Access High-Value Social Science Data</a> webpage.
include-in-header: ../../../_includes/feedback_popup.html
bibliography: BibTeX/references.bib
csl: ../../../Code/nature.csl
---

```{r}
#| echo: false

# Load in the functions.
source('../../../Code/PubMed Search_Functions.R')
```

<small>Updated: April 29<sup>th</sup>, 2026</small>

## Overview

Yale's [Data-Intensive Social Science Center (DISSC)](https://dissc.yale.edu/) was founded in 2022 to address a growing need for coordinated support for data-intensive social science research. DISSC serving as a central research support hub, expanding access to computational resources, data science tools, and statistical methods for social science researchers across the university\ [@dissc_about].

They assist colleagues with data acquisition, access, and storage, and maintain a listing of new data licenses acquired for the Yale community or publicly available in partnership with the Yale Library. Their site also includes links to additional data sources made available through other groups and sources. Their [Discover and Access High-Value Social Science Data](https://dissc.yale.edu/data-resources) webpage includes links to additional data sources made available through other groups, alongside a full listing of DISSC-managed data acquisitions. At the time of this writing, current offerings include\ [@dissc_curated_data]:

::: {style="margin-left: 2rem"}
- **Census Data:** Individual-level and household microdata from US decennial census records (1870–1950), available through Integrated Public Microdata Series (IPUMS).

- **Consumer and Spending Data:** Multiple datasets covering US consumer demographics, spending, and purchasing behavior, including longitudinal consumer history (Infutor), retail and product-level sales data (Kilts-Nielsen), de-identified credit and debit card transactions since 2019 (MBHS<sub>3</sub>), and multi-channel consumer panel data covering both e-commerce and in-store purchases since 2017 (Numerator).

- **Health and Pharmacy Data:** Drug formulary and prior authorization data covering pharmacy and medical benefit coverage across thousands of US health plans, pharmacy benefit managers, and payers, spanning 2017–2022, from Managed Markets Insight & Technology (MMIT).
:::

## Gaining Access

::: {.panel-tabset}

### Yale License

#### Do I Qualify?

DISSC advertises newly released publicly available data and maintains a curated listing of licensed datasets acquired for the Yale community. Access to some datasets may be restricted based on discipline or intended use, with restrictions varying across datasets.

#### Typical Timeline

Licenses have already been established for qualifying researchers. Review and approval processes for gaining access may vary depending on the researcher's intended use and current discipline.

#### Step-by-Step Guide

Steps to gaining access vary across DISSC-curated datasets and depend on the researcher's intended use and discipline. Please review the documentation on DISSC's data resources site or contact DISSC directly for further guidance.

:::

## Relevant Links

::: {style="margin-left: 2rem"}
- [Data-Intensive Social Science Center (DISSC)](https://dissc.yale.edu/): Find information about DISSC's resources and services.

- [Discover and Access High-Value Social Science Data](https://dissc.yale.edu/data-resources): Find DISSC's curated datasets acquired for social science research\ [@dissc_curated_data].
:::


## Publications

```{r}
#| echo: false

# Fetch and format the citations.
citations <- get_citations(
  doi = c("10.1093/ntr/ntac209"),
  affiliation = c("Yale School of Public Health", "Yale School of Medicine", "Yale University"),
  highlight_names = c("Bhramar Mukherjee")
)

# Condition the output based on presence of articles.
citations_html <- make_pub_list(citations)

if(!any(str_detect(citations$citations, "No publications"))) {
  citation_length <- length(citations$citations)
  citation_date   <- str_match(citations$query, "\\((\\d{4}):(\\d{4})\\[dp\\]\\)")[,2]
} else {
  citation_length <- 0
  citation_date   <- as.numeric(format(Sys.Date(), "%Y")) - as.numeric(str_extract("No publications found in the last 25 years.", "[0-9]{2}"))
}

```


This section presents a selection of PubMed articles that utilize one of the curated datasets and are authored by individuals affiliated with the Yale University. These articles are provided to inspire researchers and students to use the data in their own work.

::: {style="margin-left: 1rem"}

```{r}
#| echo: false
#| results: 'asis'

# Display formatted citations.
cat(citations_html)
```

```{=html}
<!-- Pagination buttons and scripting -->
<div class="citations-navigation">
  <button onclick="prevPage()">&#9664; Previous</button>
  <button onclick="nextPage()">Next &#9654;</button>
</div>

<script src="../../../../Code/pagination.js"></script>
```

:::

## References
