---
  title: "NIH's All of Us"
  description: |
         </br>
         All of Us' (AoU) one-of-a-kind dataset is stored on the Researcher Workbench, a secure, cloud-based platform. Registered researchers can access data from surveys, genomic analyses, electronic health records, physical measurements, and wearables to study the full range of factors that influence health and disease.
  title-block-banner: true
  format: html
  filters:
    - parse-latex
  markdown: 
  editor_options: 
  chunk_output_type: console
---

```{r}
#| echo: false

# Import necessary libraries
suppressPackageStartupMessages({
  library("stringr")
  library("htmltools")
  library("glue")
  library("rentrez")
  library("xml2")
  library("clipr")
})

"%!in%" <- function(x,y)!("%in%"(x,y))

# Load in the functions.
source('./PubMed_Search_Functions.R')
```

## Publications

```{r}
#| echo: false

# Fetch and format the citations.
citations <- get_citations(
  keywords = c("All of Us"), 
  affiliation = "Yale School of Public Health")

# Condition the output based on presence of articles.
citations_html <- make_pub_list(citations)
if(!any(str_detect(citations$citations, "No publications"))) {
  citation_length <- length(citations$citations)
  citation_date   <- str_match(citations$query, "\\((\\d{4}):(\\d{4})\\[dp\\]\\)")[,2]
} else {
  citation_length <- 0
  citation_date   <- as.numeric(format(Sys.Date(), "%Y")) - as.numeric(str_extract("No publications found in the last 25 years.", "[0-9]{2}"))
}

```


This section displays PubMed articles that mention the dataset and are authored by individuals affiliated with the Yale School of Public Health. Please note that the PubMed API does not distinguish between uppercase and lowercase characters when processing queries. As a result, there may be occasional mismatches in the search results.

`r citation_length` citations released after `r citation_date` were found.

::: {style="margin-left: 1rem"}

```{r}
#| echo: false
#| results: 'asis'

# Display formatted citations.
cat(citations_html)
```

```{=html}
<!-- Pagination buttons and scripting -->
<div class="citations-navigation">
  <button onclick="prevPage()">&#9664; Previous</button>
  <button onclick="nextPage()">Next &#9654;</button>
</div>

<script>
  document.addEventListener("DOMContentLoaded", function() {
    let currentPage = 1;
    const itemsPerPage = 3;
    const citationsContent = document.getElementById("citations-content");

    if (citationsContent) {
      const numPages = Math.ceil(citationsContent.children.length / itemsPerPage);

      function updateCitations() {
        const items = citationsContent.children;
        for (let i = 0; i < items.length; i++) {
          items[i].style.display = (i >= (currentPage - 1) * itemsPerPage && i < currentPage * itemsPerPage) ? "list-item" : "none";
        }
      }

      function nextPage() {
        if (currentPage < numPages) {
          currentPage++;
          updateCitations();
        }
      }

      function prevPage() {
        if (currentPage > 1) {
          currentPage--;
          updateCitations();
        }
      }

      // Initial call to update citations on load
      updateCitations();

      // Attach functions to window object for button actions
      window.nextPage = nextPage;
      window.prevPage = prevPage;
    }
  });
</script>
```

::: {.callout-note collapse="true"}
The search PubMed query used to generate this list is:

`r citations$query`
:::

:::

\
\
\
Updated: June 15th, 2025
