---
  title: "YNHHS Epic Cosmos"
  description: |
         </br>
         The Cosmos data set combines billions of clinical data points, made up of data from across the Epic community, to form the largest database of Electronic Health Record (EHR) patient information. Participating organizations provide a HIPAA-defined limited data set to the centralized Cosmos database. 
         </br>
         </br>
         With over 34,200 Clinics and 1,522 Hospitals participating in Cosmos, researchers can access a representative sample of patients who seek healthcare. You can learn more about the latest features of Cosmos on <a href="https://signin.epic.com/adfs/ls/?wa=wsignin1.0&wtrealm=https%3a%2f%2fuserweb.epic.com%2f&wctx=rm%3d1%26id%3dpassive%26ru%3d%252F&wct=2025-04-16T00%3a32%3a32Z&wreply=https%3a%2f%2fuserweb.epic.com%2f">Epic UserWeb</a>, available to all Yale New Haven Health System (YNHHS) research users and on the <a href="https://teams.microsoft.com/l/team/19%3AJTlr92eAMEnJsBJmV1Uymbla8oek8ePqASKzxPE7mrQ1%40thread.tacv2/conversations?groupId=aa0ac073-1651-4dee-a59e-884bef65ea6b&tenantId=dd8cbebb-2139-4df8-b411-4e3e87abeb5c">Yale Cosmos Teams</a> Site.
  title-block-banner: true
  format: html
  filters:
    - parse-latex
  markdown: 
  editor_options: 
  chunk_output_type: console
---

```{r}
#| echo: false

# Import necessary libraries
suppressPackageStartupMessages({
  library("stringr")
  library("htmltools")
  library("glue")
  library("rentrez")
  library("xml2")
  library("clipr")
})

"%!in%" <- function(x,y)!("%in%"(x,y))

# Load in the functions.
source('./PubMed_Search_Functions.R')
```

## Overview

Researchers wanting to use the Yale New Haven Health System (YNHHS) Epic and Epic Cosmos data will need a Researcher Basic Access (RBA) account. Qualifying researchers will receive a YNHHS NetID and Duo Access, which affords access to:

:::::: columns
::: {.column style="text-align:right; width:25%"}
[Virtual Desktop Infrastructure (VDI)](https://medicine.yale.edu/ybic/computational-resources/ynhhs/#virtual-desktop-infrastructure)

\
\
\
\hyperlink{https://medicine.yale.edu/ybic/data-resources-services/slicerdicer/}{Epic SlicerDicer}

\
\
\
\
\
\

\hyperlink{https://medicine.yale.edu/ybic/computational-resources/ynhhs/#computational-health-platform}{Computational Health Platform (CHP)}\
\
\
\

\
\

Microsoft Office 365
:::

::: {.column width="3%"}
<!-- empty column to create gap -->
:::

::: {.column width="72%"}
Remote access to a YNHHS secure device without requiring the YNHHS VPN. VDI includes the following software: SQL Server Management Studio, R, Stata, SAS, Office365, and Epic. It can be used to handle small-to-medium sized data, accessing Epic charts, SlicerDicer, and the Computational Health Platform (CHP).

Access de-identified summary count data representing the entire YNHHS patient dataset for data exploration, analytics, and reporting. Institutional Review Board (IRB) applications are not required to access de-identified data, but training is strongly recommended before usage. 

Researchers can request access to the Epic Cosmos system, which will provide de-identified, summary count data for all participating Epic systems outside of YNHHS.

A secure, compliant, and collaborative high-performance computing cluster specifically designed for clinical data analysis. Suitable for AI research and development, and is well suited for analyzing large volumes of data in different forms (i.e. images, text, structured).

This tool requires advanced data science skills with relational databases, Python, and SQL. Additional details can be found here: [OMOP Common Data Model](https://medicine.yale.edu/ybic/data-resources-services/omop-common-data-model/).

Microsoft applications linked within the YNHHS organization.
:::
::::::

There are three levels of access available:

::: {style="margin-left: 2rem"}
-   RBA without an IRB give access to the afore mentioned resources but only supplies de-identified, summary count data produced within YNHHS Epic.
-   RBA with an approved IRB gives access to individual-level data for Epic data produced within YNHHS that is strictly relevant to the IRB approved study.
-   Epic Cosmos - RBA holders can gain access to de-identified, summary count data produced by Epic participating institutions outside of YNHHS upon special request.
:::




### Do I Qualify?

All Yale University and YNHHS (YU-YNHHS) Covered Entity faculty and licensed provider may submit and sponsor RBA requests. If you are within the Yale system, your manager or sponser with a YNHHS account must complete the ["YNHHS Researcher Basic Access"](https://shorturl.at/nJOS0) request on your behalf to start the process.

If you are not a member of a YU-YNHHS Covered Entity (e.g. YSPH), you will need a sponsor to submit a request on your behalf. Find the current "List of Covered Departments" under "Administrative Functions" on the [hippa.yale.edu](https://hipaa.yale.edu/resources/resource-documents-and-guidance) resources page.

## Step-by-Step

::: {style="margin-left: 2rem"}
1.  Your YU-YNHHS Covered Entity manager or sponsor will complete the ["YNHHS Researcher Basic Access"](https://shorturl.at/nJOS0) request on your behalf. This will give you YNHHS log-in credentials and identification authentication with Duo.

2.  **For Yale NetID holders:** [HIPAA Privacy and Security](https://hipaa.yale.edu/training/training-modules) courses in Workday, including the basic privacy and security module followed by the role specific module (i.e. for clinician's or researcher's).

    **For Non-Yale NetID holders:** read and sign the [HIPAA Privacy and Security](https://hipaa.yale.edu/training/training-modules) training materials specific to your role on the project (i.e. for clinician's or researcher's). Submission directions for these documents can be found on the PDF.

3.  Human Subjects Protection Training (HSPT) certificate from CITI that has been completed within the last 3 years. Below is a summary of the process and the complete directions can be found on the [It's Your Yale](https://your.yale.edu/research-support/human-research-protection-program/education-and-training/human-research-training) page about HSPT. You can reach out to [irb.training@yale.edu](mailto:irb.training@yale.edu) for further questions.

    Researchers will need to complete the following CITI certificates:

::: {style="margin-left: 2rem"}
-   "Basic Training"
-   One of the following two initial/basic Human Subjects Research modules: "Biomedical Research Basic" or "Social Behavioral Educational Researchers Basic".
:::

::: {style="margin-left: 0rem"}
For Yale-affiliates, it is preferable, though not required, to go through Workday to access the CITI training modules. Doing so will expedite recognition of the certificate in the Yale Integrated Research Enterprise Solution (IRES) IRB system. In Workday, the CITI training link will show up as the "Human Subjects Protection Training – Initial/Refresher" course. You can reach out to your manager, [learning@yale.edu](mainto:learning@yale.edu), or [irb.training@yale.edu](mainto:irb.training@yale.edu) to request manual assignment if needed.

Qualifying certificates completed through other institutions CITI subscriptions need to be emailed to [irb.training@yale.edu](mainto:irb.training@yale.edu). Please keep in mind that receiving CITI certification independently from a subscribing institution will incur charges.
:::

4.  SlicerDicer

::: {style="margin-left: 2rem"}
- https://jdat.ynhh.org/AnalyticsPortal/helix
:::

:::

## Typical Timeline

::: callout-important
If you are looking to gain access to the individual-level YNHHS Epic data you will need to submit an IRB application. Time-to access in these cases is contingent upon the IRB's process that will vary for applications.
:::

::: {.panel-tabset}

## YNHHS Epic

Time to get RBA to YNHHS de-identified data.

## Epic Cosmos

Time to get access to de0identified data from all Epic centers outside YNHHS.

:::


## Important Links

::: {style="margin-left: 2rem"}
- [YNHHS Learning Management System (LMS)](https://ynhh.certpointsystems.com/): Center Point is where you can sign up for in-person or online training courses.
- [MyAPPs](https://myapps.ynhh.org): Access to the VDI via a Citrix Workspace download. YNHHS DUO and YU/YNHHS VPN is required if not accessing on the YU or YNHHS network.

:::

## Publications

```{r}
#| echo: false

# Fetch and format the citations.
citations <- get_citations(
  keywords = c("Yale New Haven Health System Epic", "Yale New Haven Health System Epic Cosmos", "YNHHS Epic", "YNHHS Epic Cosmos"), 
  affiliation = "Yale School of Public Health")

# Condition the output based on presence of articles.
citations_html <- make_pub_list(citations)
if(!any(str_detect(citations$citations, "No publications"))) {
  citation_length <- length(citations$citations)
  citation_date   <- str_match(citations$query, "\\((\\d{4}):(\\d{4})\\[dp\\]\\)")[,2]
} else {
  citation_length <- 0
  citation_date   <- as.numeric(format(Sys.Date(), "%Y")) - as.numeric(str_extract("No publications found in the last 25 years.", "[0-9]{2}"))
}

```


This section displays PubMed articles that mention the dataset and are authored by individuals affiliated with the Yale School of Public Health. Please note that the PubMed API does not distinguish between uppercase and lowercase characters when processing queries. As a result, there may be occasional mismatches in the search results.

`r citation_length` citations released after `r citation_date` were found.

::: {style="margin-left: 1rem"}

```{r}
#| echo: false
#| results: 'asis'

# Display formatted citations.
cat(citations_html)
```

```{=html}
<!-- Pagination buttons and scripting -->
<div class="citations-navigation">
  <button onclick="prevPage()">&#9664; Previous</button>
  <button onclick="nextPage()">Next &#9654;</button>
</div>

<script>
  document.addEventListener("DOMContentLoaded", function() {
    let currentPage = 1;
    const itemsPerPage = 3;
    const citationsContent = document.getElementById("citations-content");

    if (citationsContent) {
      const numPages = Math.ceil(citationsContent.children.length / itemsPerPage);

      function updateCitations() {
        const items = citationsContent.children;
        for (let i = 0; i < items.length; i++) {
          items[i].style.display = (i >= (currentPage - 1) * itemsPerPage && i < currentPage * itemsPerPage) ? "list-item" : "none";
        }
      }

      function nextPage() {
        if (currentPage < numPages) {
          currentPage++;
          updateCitations();
        }
      }

      function prevPage() {
        if (currentPage > 1) {
          currentPage--;
          updateCitations();
        }
      }

      // Initial call to update citations on load
      updateCitations();

      // Attach functions to window object for button actions
      window.nextPage = nextPage;
      window.prevPage = prevPage;
    }
  });
</script>
```

::: {.callout-note collapse="true"}
The search PubMed query used to generate this list is:

`r citations$query`
:::

:::


## References

The content provided here is a compilation of work from multiple groups throughout YU and YNHHS:

::: {style="margin-left: 2rem"}
1.  [Yale Biomedical Informatics & Computing](https://medicine.yale.edu/ybic)

:::

\
\
\
Updated: June 15th, 2025
