---
  title: "National Health and Nutrition Examination Survey"
  description: |
         </br>
         The National Health and Nutrition Examination Survey (NHANES) collects data about the health of adults and children in the United States. We also collect data about what participants eat, drink, and take as supplements to determine how many nutrients are in their diet. These dietary interviews and blood tests help us measure the nutritional status of U.S. adults and children\ [@about_nhanes; @achievements_nhanes].
         </br>
         </br>
         CDC's <a href="https://www.cdc.gov/nchs/index.html">National Center for Health Statistics</a> (NCHS) conducts NHANES. NHANES is the only national health survey that includes health exams, laboratory tests, and dietary interviews for participants of all ages.
         </br>
         </br>
         NHANES data can help improve the health of Americans. Survey data have driven changes in how doctors treat patients and how public policy supports good health.
  title-block-banner: true
  format: html
  header-includes:
    - '<link href="https://cdnjs.cloudflare.com/ajax/libs/font-awesome/5.15.3/css/all.min.css" rel="stylesheet">'
  editor: source
  bibliography: BibTeX/nhanes_references.bib
  csl: BibTeX/nature.csl
  filters:
    - parse-latex
  markdown: 
  editor_options: 
  chunk_output_type: console
---

```{r}
#| echo: false

# Import necessary libraries
suppressPackageStartupMessages({
  library("stringr")
  library("htmltools")
  library("glue")
  library("rentrez")
  library("xml2")
  library("clipr")
})

"%!in%" <- function(x,y)!("%in%"(x,y))

# Load in the functions.
source('./PubMed_Search_Functions.R')
```

Updated: August 10th, 2025

## Overview

Surveillance of the dietary habits of the United States population began with the passage of the National Health Survey Act in 1956. Initially, the focus was on statistically measuring the extent, distribution, and impact of chronic illnesses, such as cardiovascular disease and diabetes, and disabilities through three National Health Examination Surveys (NHES). The latter two cycles of NHES focused on the growth and development of children and adolescents. In the early 1970s, this scope was expanded to include research into the connections between diet and diseases with the addition of the National Nutritional Surveillance System to NHES, forming NHANES\ [@Zipf2013; @history_nhanes].

This evolution included incorporating measurements that tracked changes in nutritional status over time, alongside clinical observations and professional assessments of survey participants. The NHANES survey underwent four changes until the protocol was settled in 1999 on the continuously conducted surveillance that is still carried out today\ [@history_nhanes].

## Programs

Through its different iterations, NHANES, and previously NHES from 1960 to 1970, collected various surveys and metrics from differently sized and demographic samplings of the U.S. population. A researcher seeking to use this data will need to keep in mind that each survey iteration was conducted with specific differences, where some are more cross-comparable than others. We briefly discuss these differences here, but additional methadological variations can be found by reviewing the references provided below.

### Methods and Representational Differences

The following table summarizes critical program features and highlights changes or differences between each program iteration. Although the data types, such as interviews and physician examinations, have remained consistent since the program's inception, the kinds and scope of exams and tests have changed. The base population also varied, with certain programs intentionally designed to oversample underrepresented ethnic and racial groups or vulnerable populations to improve research possibilities for these groups.

We do not spend time distinguishing methodological differences if the same type of exam was conducted with different instruments or protocols. Researchers interested in learning more about these granular differences to assess the possibility of cross-program compatibility are recommended to review the [Historical NHANES Component Matrix](https://wwwn.cdc.gov/nchs/data/nhanes/Historical_NHANES_component_matrix.pdf), which compares NHANES programs up to 2006, and the [National Health and Nutrition Examination Survey: 1999–2022 Survey Content Brochure](https://wwwn.cdc.gov/nchs/data/nhanes/survey_contents.pdf), which summarizes the annual continuous NHANES protocol differences. Additional information is also provided in the respective programs' _Survey Methods and Analytic Guidelines_ documentation or other manuals, accessible on the CDC landing page for each program\ [@historic_nhanes_content; @continuous_brochure].

<table class="custom-table">
    <thead>
        <tr>
            <th class="toggle"></th>
            <th class="iteration">Iteration</th>
            <th class="design">Design and Data</th>
            <th class="var">Variable Categories</th>
            <th class="population">Population</th>
        </tr>
    </thead>
    <tbody>
        <!-- First Row -->
        <tr id="row1a">
            <td class="toggle">
              <button class="toggle-btn" data-target="row1a">
                <i class="fa fa-angle-down"></i>
              </button>
            </td>
            <td class="iteration"><a href="https://wwwn.cdc.gov/nchs/nhanes/nhes1/default.aspx" target="_blank">NHES I</a>\ [@nhes_I]</td>
            <td class="design">
              <span class="hidden-content hidden">
                <strong>Dates:</strong> 1959-62
                </br>
                </br>
                <strong>Study Design:</strong> A multistage, stratified probability sample of loose clusters of individuals in land-based segments was used. Participants were randomly selected from primary sampling units (PSUs) that were partitioned into segments based on geography, eligibility, and housing density\ [@Linder1965; @nhne_design].
                </br>
                </br>
                <strong>Data Types:</strong>\ [@Linder1965]
                <ul>
                  <li>Household interview</li>
                  <li>Available health records</li>
                  <li>Direct medical and dental examination (test results and measurements)</li>
                </ul>
                </br>
                <strong>Data Files:</strong> TXT, SAS, and PDF
              </span>
            </td>
            <td class="var">
              <span class="hidden-content hidden">
                <ul>
                  <li>Demographic</li>
                  <li>Socioeconomic</li>
                  <li>Geographic (state and urbanicity)</li>
                  <li>Sampling weights to reflect 111 million persons of the base population</li>
                  <li>Psychological distress</li>
                  <li>Physical measurements (18 measurements of body size)</li>
                  <li>Cardiovascular data</li>
                  <li>Osteoarthritis and rheumatoid arthritis</li>
                  <li>Dental findings</li>
                  <li>Diabetes findings</li>
                  <li>Vision data</li>
                </ul>
              </span>
            </td>
            <td>
              <span class="hidden-content hidden">
                <ul>
                  <li>7,800 were contacted with 7,710 interviewed and, of these, 6,672 examined</li>
                  <li>18-79 years of age</li>
                  <li>Noninstitutionalized civilian population</li>
                </ul>
              </span>
            </td>
        </tr>
        <!-- Second Row -->
        <tr id="row2a">
          <td class="toggle">
            <button class="toggle-btn" data-target="row2a"><i class="fa fa-angle-down"></i></button>
          </td>
          <td class="iteration"><a href="https://wwwn.cdc.gov/nchs/nhanes/nhes2/default.aspx" target="_blank">NHES II</a>\ [@nhes_II]</td>
          <td class="design">
            <span class="hidden-content hidden">
              <strong>Dates:</strong> 1963-65
              </br>
              </br>
              <strong>Study Design:</strong> A similar strategy to NHES I was employed\ [@Secretary1967].
              </br>
              </br>
              <strong>Data Types:</strong>\ [@Linder1965; @Secretary1967]
                <ul>
                  <li>Household interview with the household adult, usually the mother</li>
                  <li>Available health records provided by the household adult</li>
                  <li>School questionnaire</li>
                  <li>Direct medical and dental examination (test results and measurements)</li>
                  <li>Vision and hearing tests</li>
                  <li>Anthropometric measurements</li>
                </ul>
              </br>
              <strong>Data Files:</strong> TXT, SAS, and PDF
            </span>
          </td>
          <td class="var">
            <span class="hidden-content hidden">
              <ul>
                <li>Demographic (including twin status)</li>
                <li>Socioeconomic</li>
                <li>Family characteristics</li>
                <li>Geographic (state and urbanicity)</li>
                <li>Sampling weights to reflect 24 million children of the base population</li>
                <li>Medical and developmental history</li>
                <li>School questionnaire (i.e. academic standing, use of special resources)</li>
                <li>Psychological examination</li>
                <li>Physical examination</li>
                <li>Ear, nose, and throat examination</li>
                <li>Diagnostic impressions (10 records including neurological)</li>
                <li>Audiometry</li>
                <li>Vision</li>
                <li>Dental findings</li>
                <li>Body measurements (13 sitting and 13 standing)</li>
                <li>Bone age by X-ray</li>
              </ul>
            <span class="hidden-content hidden">
          </td>
          <td>
            <span class="hidden-content hidden">
              <ul>
                <li>7,417 interviewed with 7,119 examined</li>
                <li>6-11 years of age</li>
                <li>Parents from the noninstitutionalized civilian population</li>
                <li>Sampling was restricted to the conterminous United States.</li>
              </ul>
            </span>
          </td>
        </tr>
        <!-- Third Row -->
        <tr id="row3a">
          <td class="toggle">
            <button class="toggle-btn" data-target="row3a"><i class="fa fa-angle-down"></i></button>
          </td>
          <td class="iteration"><a href="https://wwwn.cdc.gov/nchs/nhanes/nhes3/default.aspx" target="_blank">NHES III</a>\ [@nhes_III]</td>
          <td class="design">
            <span class="hidden-content hidden">
              <strong>Dates:</strong> 1966-70
              </br>
              </br>
              The study design, data types, and available files types are the same as those of NHES II\ [@Woolsey1974].
            </span>
          </td>
          <td>
            <span class="hidden-content hidden">
              Variables tracked are the same as NHES II.
            </span>
          </td>
          <td>
            <span class="hidden-content hidden">
              <ul>
                <li>7,514 interviewed with 6,768 examined</li>
                <li>12-17 years of age</li>
                <li>Parents from the noninstitutionalized civilian population</li>
                <li>Sampling was restricted to the conterminous United States.</li>
              </ul>
              </br>
              <strong>NOTE:</strong> Nearly 1/3$^{rd}$ of the children examined in NHES II were reexamined in NHES III. The Extended Data Tape (EDT) helps the user identify the entries present in both cycles to create a composite data set for longitudinal studies.
            </span>
          </td>
        </tr>
        <!-- Fourth Row -->
        <tr id="row4a">
          <td class="toggle">
            <button class="toggle-btn" data-target="row4a"><i class="fa fa-angle-down"></i></button>
          </td>
          <td class="iteration"><a href="https://wwwn.cdc.gov/nchs/nhanes/nhanes1/default.aspx" target="_blank">NHANES I</a>\ [@nhanes_I]</td>
          <td class="design">
            <span class="hidden-content hidden">
              <strong>Dates:</strong> 1971-74
              </br>
              </br>
              <strong>Study Design:</strong> Samples of persons were selected through complex, multistage, highly stratified, cluster samples of households. Similarly to NHES, participants were randomly selected from primary sampling units (PSUs) that were partitioned into segments based on geography, eligibility, and housing density\ [@Linder1965].
              </br>
              </br>
              <strong>Data Types:</strong>\ [@Harvey1981]
              </br>
              The following was collected from all participants
              <ul>
                <li>Household interview</li>
                <li>Available health records and history</li>
                <li>Direct medical, dental, and dermatological examination (test results and measurements)</li>
                <li>Vision and hearing tests</li>
                <li>Anthropometric measurements</li>
                <li>Hand-wrist X-rays for participants ages 1-17</li>
                <li>Nutrient composition, frequency, and adequacy</li>
              </ul>
              </br>
              Detailed examinations included the above list and
              <ul>
                <li>Supplement medical history for arthritis, respiratory and cardiovascular conditions</li>
                <li>Health care needs questionnaire</li>
                <li>General well-being questionnaire</li>
                <li>Extended examination and tests</li>
                <li>X-rays of the chest and hip and knee joints</li>
              </ul>
              </br>
              <strong>Data Files:</strong> TXT, SAS, or XPT and PDF
            </span>
          </td>
          <td class="var">
            <span class="hidden-content hidden">
              Additional details concerning the two baseline populations are summarized in the [_Plan and Operation of the Health and Nutrition Examination Survey Series 1 Number 10a_](https://www.cdc.gov/nchs/data/series/sr_01/sr01_010a.pdf) PDF pg. 20.
              </br>
              </br>
              <ul>
                <li>Demographic</li>
                <li>Socioeconomic</li>
                <li>Family characteristics</li>
                <li>Geographic (state and urbanicity)</li>
                <li>Sampling weights to reflect the population of 1-74 year olds within 60 age-sex-race groups at the midpoint of the survey</li>
                <li>Medical history questionnaire, ages 1-11 years,</li>
                <li>Medical history questionnaire, ages 12-74 years</li>
                <li>Medical examination, ages 1-74 years</li>
                <li>Health care needs, general medical history supplement, and respiratory and cardiovascular supplements, ages 25-74 years</li>
                <li>General well-being and the CES-D depression scale developed by The National Institute of Mental Health, ages 25-74 years</li>
                <li>Dental, ages 1-74 years</li>
                <li>Dermatology, ages 1-74 years</li>
                <li>Near and distant vision, ages 25-74 years</li>
                <li>Ophthalmology, ages 1-74 years</li>
                <li>Audiometric</li>
                <li>Spirometry-best trials only, ages 25-74 years</li>
                <li>Anthropometry goniometry, skeletal age, bone density, and cortical thickness, Ages 1-74 years</li>
                <li>Arthritis</li>
                <li>Chest X-ray, pulmonary diffusion, and tuberculin test results, ages 25-74 years</li>
                <li>Electrocardiogram</li>
                <li>Computer measurements and interpretations of electrocardiograms, ages 25-74 years</li>
                <li>Biochemistry, serology, hematology, blood slides, urine dipstick</li>
                <li>Model gram</li>
                <li>Nutrient composition</li>
                <li>Dietary frequency and adequacy, ages 1-74 years</li>
                <li>24-hour food consumption intake, ages 1-74 years</li>
              </ul>
            </span>
          </td>
          <td>
            <span class="hidden-content hidden">
              <ul>
                <li>32,000 were contacted with 31,973 interviewed and, of these, 23,808 examined</li>
                <li>Noninstitutionalized civilian population</li>
                <li>Sampling was restricted to the conterminous United States and excluded individuals residing in Native American reservations.</li>
              </ul>
              </br>
              The base population surveyed and examined was divided into two groups of interest: one primarily focused on nutrition, and a subset of this group that underwent a more detailed process. Each of these had different sampling emphases, which is summarized below\ [@Rice1973; @nhanes_I].
              </br>
              </br>
              <strong>Standard examination cohort:</strong>
              </br>
              <ul>
                <li>1-74 years of age</li>
                <li>Emphasis by oversampling of the following groups: at or below the federal poverty line, preschool children (6 months - 5 years), women of child-bearing age, and persons aged 60-74 years old.</li>
              </ul>
              </br>
              <strong>Detailed examination cohort:</strong>
              </br>
              <ul>
                <li>6,913 of the 32,000. 3,059 received detailed medical examinations during the survey and an additional 3,059 received detailed examinations after completion of the nutrition survey in October 1975. The last group is referred to as the NHANES I Augmentation group.</li>
                <li>25-74 years of age</li>
                <li>No oversampling relative to the national population</li>
              </ul>
            </span>
          </td>
        </tr>
        <!-- Fifth Row -->
        <tr id="row5a">
          <td class="toggle">
            <button class="toggle-btn" data-target="row5a"><i class="fa fa-angle-down"></i></button>
          </td>
          <td class="iteration"><a href="https://wwwn.cdc.gov/nchs/nhanes/nhanes2/default.aspx" target="_blank">NHANES II</a>\ [@nhanes_II]</td>
          <td class="design">
            <span class="hidden-content hidden">
              <strong>Dates:</strong> 1976-80
              </br>
              </br>
              <strong>Study Design:</strong> A similar strategy to NHANES I was employed\ [@Rice1981].
              </br>
              </br>
              <strong>Data Types:</strong>\ [@Rice1981]
              </br>
              <ul>
                <li>Household interview</li>
                <li>Household examination to allow inclusion of very young children and the elderly</li>
                <li>Available health records and history</li>
                <li>Medications and vitamin usage</li>
                <li>Behavior questionnaire for 25-74 year olds</li>
                <li>Direct medical, dental, and dermatological examination (test results and measurements)</li>
                <li>Hearing tests on 4-19 year olds and speech tests on 4-6 year olds</li>
                <li>Anthropometric measurements</li>
                <li>Cervical and lumbar spine X-rays (excluding women under 50 years of age) and chest X-rays. All X-rays were done on participants 25-74 years of age and excluded pregnant women</li>
                <li>Nutrient composition, frequency, and adequacy</li>
              </ul>
              </br>
              <strong>Data Files:</strong> TXT, SAS, or XPT and PDF
            </span>
          </td>
          <td class="var">
            <span class="hidden-content hidden">
              <ul>
                <li>Demographic</li>
                <li>Socioeconomic</li>
                <li>Family characteristics</li>
                <li>Geographic (state and urbanicity)</li>
                <li>Sampling weights to reflect the population of 6 months - 74 year olds within 76 age-sex-income groups at the midpoint of the survey</li>
                <li>Medical history questionnaire, ages 6 months-11 years</li>
                <li>Medical history questionnaire, ages 12-74 years</li>
                <li>Health history supplement, ages 12-74 years</li>
                <li>Physician's examination</li>
                <li>Anthropometry</li>
                <li>Audiometric air conduction test, ages 4-19 years</li>
                <li>Behavior questionnaire, ages 25-74</li>
                <li>Chest X-ray examination</li>
                <li>Electrocardiogram</li>
                <li>Hematology and biochemistry, ages 6 months-74 years</li>
                <li>Hepatitis A and B (Updated October 2007)</li>
                <li>Herpes I and II (Updated September 2016)</li>
                <li>Syphilis serological markers</li>
                <li>Allergy skin testing</li>
                <li>Model gram</li>
                <li>Nutrient composition</li>
                <li>Total nutrient intake, food frequency, and other related dietary data</li>
                <li>24-hour recall, specific food item</li>
              </ul>
            </span>
          </td>
          <td>
            <span class="hidden-content hidden">
              <ul>
                <li>27,801 were contacted with 25,286 interviewed and, of these, 20,322 examined</li>
                <li>6 months - 74 years of age</li>
                <li>Noninstitutionalized civilian population</li>
                <li>All 50 States</li>
                <li>Emphasis by oversampling of the following groups: at or below the federal poverty line, preschool children (6 months - 5 years), and persons aged 60-74 years old</li>
              </ul>
              </br>
              <strong>NOTE:</strong> Efforts were made to maintain the variables and procedures conducted in NHANES I, but not everything was kept the same. Refer to the procedure documentation for clarification on opportunities for longitudinal evaluation\ [@Rice1981; @historic_nhanes_content].
            </span>
          </td>
        </tr>
        <!-- Sixth Row -->
        <tr id="row6a">
          <td class="toggle">
            <button class="toggle-btn" data-target="row6a"><i class="fa fa-angle-down"></i></button>
          </td>
          <td class="iteration"><a href="https://wwwn.cdc.gov/nchs/nhanes/hhanes/default.aspx" target="_blank">HANES</a>\ [@hhanes]</td>
          <td class="design">
            <span class="hidden-content hidden">
              <strong>Dates:</strong> 1982-84
              </br>
              </br>
              <strong>Study Design:</strong> A similar strategy to NHANES I was employed, with one difference. Instead of determining the primary sampling units (PSUs) by partitioning the national map, only specific regions where the subgroup populations were most represented were included\ [@Feinleib1985].
              </br>
              </br>
              <strong>Data Types:</strong>\ [@Rice1981; @Feinleib1985]
              </br>
              <ul>
                <li>Household interview</li>
                <li>Household examination to allow inclusion of very young children and the elderly</li>
                <li>Available health records and history</li>
                <li>Alcohol consumption and drug abuse</li>
                <li>Measures of depression, ages 20-74 Years</li>
                <li>Direct medical, dental, and vision examination (test results and measurements)</li>
                <li>Hearing, ages 6 months - 74 years</li>
                <li>Anthropometric measurements</li>
                <li>Nutrient composition, frequency, and adequacy</li>
                <li>Pesticide and lead exposure, ages 20-74 years</li>
              </ul>
              </br>
              <strong>Data Files:</strong> TXT, SAS, or XPT and PDF
            </span>
          </td>
          <td class="var">
            <span class="hidden-content hidden">
              <ul>
                <li>Demographic</li>
                <li>Socioeconomic</li>
                <li>Family characteristics</li>
                <li>Geographic (state and urbanicity)</li>
                <li>Sampling weights to reflect the population of 6 months - 74 year olds within each subgroup</li>
                <li>Adolescent and adult history questionnaire, ages 12-74 years</li>
                <li>Child history questionnaire, ages 6 months - 11 years</li>
                <li>Physician's examination, ages 6 months - 74 years</li>
                <li>Body measurements</li>
                <li>Dental and hearing health, ages 6 months - 74 years</li>
                <li>Vision, ages 6-74 years</li>
                <li>Alcohol consumption and drug abuse, ages 12-74 years</li>
                <li>Blood and urine assessments, ages 6 months - 74 years</li>
                <li>Pesticide data, ages 20-74 years</li>
                <li>Gallbladder ultrasound data, ages 20-74 years</li>
                <li>Diabetes and OGTT data, ages 20-74 years</li>
                <li>Dietary practices, food frequency and total nutrient intake</li>
                <li>24-Hour recall, ages 6 months - 74 years</li>
                <li>Measurement and interpretation of electrocardiograms, ages 20-74 years</li>
                <li>Measures of depression, ages 20-74 years</li>
              </ul>
            </span>
          </td>
          <td>
            <span class="hidden-content hidden">
              <ul>
                <li>16,000 were contacted with 13,689 interviewed and, of these, 11,653 examined\ [@Feinleib1985; @Zipf2013]</li>
                <li>Of those examined: 7,462 Mexican-American, 1,357 Cuban-American, and 2,834 Puerto Rican\ [@Zipf2013]</li>
                <li>6 months - 74 years</li>
                <li>Noninstitutionalized civilian population</li>
                <li>"Eligible" Hispanic subgroups from specified regions of the United States included Mexican-Americans in the five Southwestern states (CA, AZ, NM, TX, and CO), Puerto Ricans in the New York City area (defined as selected counties in New York, New Jersey, and Connecticut), and Cuban-Americans in Dade County (Miami), Florida.</li>
                <li>Emphasis by oversampling of the following groups: 6 months - 19 years of age and 45–74 years of age</li>
              </ul>
              </br>
              <strong>NOTE:</strong> Efforts were made to maintain the variables and procedures conducted in NHANES I and II, but not everything was kept the same. Refer to the procedure documentation for clarification on opportunities for longitudinal evaluation\ [@Feinleib1985; @historic_nhanes_content].
            </span>
          </td>
        </tr>
        <!-- Seventh Row -->
        <tr id="row7a">
          <td class="toggle">
            <button class="toggle-btn" data-target="row7a"><i class="fa fa-angle-down"></i></button>
          </td>
          <td class="iteration"><a href="https://wwwn.cdc.gov/nchs/nhanes/nhanes3/default.aspx" target="_blank">NHANES III</a>\ [@nhanes_III]</td>
          <td class="design">
            <span class="hidden-content hidden">
              <strong>Dates:</strong> 1988-94
              </br>
              </br>
              <strong>Study Design:</strong> A similar strategy to NHANES I was employed\ [@Feindleib1994].
              </br>
              </br>
              <strong>Data Types:</strong>\ [@Rice1981; @Feindleib1994]
              </br>
              <ul>
                <li>Household interview</li>
                <li>Household examination to allow inclusion of very young children and the elderly</li>
                <li>Available health records and history</li>
                <li>Alcohol consumption and drug abuse</li>
                <li>Direct medical, dental, hearing, and vision examination (test results and measurements)</li>
                <li>Anthropometric measurements</li>
                <li>Nutrient composition, frequency, adequacy, and nutritional supplementation</li>
                <li>Pesticide, volatile organic compounds (VOC’S), and lead exposure</li>
                <li>Knee X-ray's</li>
              </ul>
              </br>
              Data were collected over two phases (Phase 1 from 1988-91 and Phase 2 from 1991-94) and organized into three publicly available categories: core content, augmented/special studies (including multiple imputation and raw curve spirometry), and surplus sera results. The latter two categories were prepared after the completion of NHANES III.
              </br>
              </br>
              <strong>Data Files:</strong> TXT, SAS, or XPT and PDF
              </br>
              </br>
              <strong>NOTE:</strong> NHANES III has both publicly available and restricted data files. You can learn more about both by navigating from the primary NHANES III landing page to the [Data Files](https://wwwn.cdc.gov/nchs/nhanes/nhanes3/datafiles.aspx) subpage.
            <span>
          </td>
          <td class="var">
            <span class="hidden-content hidden">
              Core NHANES III data were broadly divided into five assessments: household interviews for adults and youths, examinations, laboratory results, and dietary recalls. You can learn more about the contents of these assessments in their respective documentation, which also includes a table cross-comparing what is included in each\ [@nhanes_III; @Feindleib1994;  @nhanes_III_adult_data_file].
              </br>
              </br>
              <ul>
                <li>Demographic</li>
                <li>Socioeconomic</li>
                <li>Family characteristics</li>
                <li>Geographic (state and urbanicity)</li>
                <li>Sampling weights to reflect the population of 2 months and older within each subgroup</li>
                <li>Social support and residence questionnaire for adults</li>
                <li>General physical examination and central nervous system evaluation</li>
                <li>Body measurements</li>
                <li>Infant feeding practices/diet</li>
                <li>School attendance and cognitive function for youths</li>
                <li>Wechsler Intelligence Scale for Children (WISC) and the Wide Range Achievement Test (WRAT)</li>
                <li>Dental, vision, and hearing questions</li>
                <li>Blood and urine assessments</li>
                <li>Respiratory and allergy questionnaire and examinations</li>
                <li>Fundus photography and bone density</li>
                <li>Dietary practices, food frequency and total nutrient intake</li>
                <li>24-Hour recall, ages 6 months - 74 years</li>
                <li>Drug abuse and alcohol use examination</li>
                <li>Tobacco use questionnaire and examination </li>
                <li>Physical activity questionnaire for adults and examination for all participants</li>
                <li>Functional impairment</li>
                <li>Environmental and occupational exposures, such as pesticides, VOCs, and lead</li>
              </ul>
            <span>
          </td>
          <td>
            <span class="hidden-content hidden">
              <ul>
                <li>39,695 were contacted with 33,994 interviewed and, of these, 30,818 examined. 493 received special, limited examinations in their home.</li>
                <li>2 months of age and older</li>
                <li>Noninstitutionalized civilian population</li>
                <li>All 50 States</li>
                <li>Emphasis by oversampling of the following groups: preschool children (2 months - 5 years), and persons aged 60 and over, Mexican-American persons, and non-Hispanic black persons\ [@nhanes_III; @nhne_design]</li>
              </ul>
            <span>
          </td>
        </tr>
        <!-- Eighth Row -->
        <tr id="row8a">
          <td class="toggle">
            <button class="toggle-btn" data-target="row8a"><i class="fa fa-angle-down"></i></button>
          </td>
          <td class="iteration"><a href="https://wwwn.cdc.gov/nchs/nhanes/default.aspx" target="_blank">Continuous NHANES</a>\ [@data_nhanes]</td>
          <td class="design">
            <span class="hidden-content hidden">
              <strong>Dates:</strong> Every two years since 1999
              </br>
              </br>
              <strong>Study Design:</strong> A similar strategy to NHANES I was employed. For each two-year cycle, approximately 30 counties out of the 3,000 PSU counties were selected, ensuring that each of the four regions of the United States, as well as both metropolitan and nonmetropolitan areas, were represented each time\ [@Zipf2013]. The notable exceptions to this continuity are the 2019-20 and 2021-23 cycles\ [@nhanes_2019_20; @Terry2021].
              </br>
              </br>
              <strong>Data Types:</strong>\ [@Rice1981; @Feindleib1994]
              </br>
              <ul>
                <li>Household interview and follow-up phone call</li>
                <li>Available health records and history</li>
                <li>Alcohol consumption and drug abuse</li>
                <li>Direct medical, dental, hearing, and vision examination (test results and measurements)</li>
                <li>Anthropometric measurements</li>
                <li>Nutrient composition, frequency, adequacy, and nutritional supplementation</li>
                <li>Environmental and occupational exposures, such as pesticides, VOCs, and lead</li>
                <li>Dual-Energy X-ray Absorpiometry (DEXA) scan</li>
              </ul>
              </br>
              <strong>Data Files:</strong> DSPI Doc or XPT
              </br>
              </br>
              <strong>NOTE:</strong> Continuous NHANES has both publicly available and restricted data files. To learn more about both, navigate to the respective data release year accessible from the main [NHANES Questionnaires, Data, and Related Documentation](https://wwwn.cdc.gov/nchs/nhanes/default.aspx) webpage.
              </br>
              </br>
              <strong>NOTE:</strong> Generally speaking, any 2-year data cycle in Continuous NHANES can be combined with adjacent 2-year data cycles utilizing the same sampling design to create analytic data files based on 4 or more years of data, allowing for greater precision. Protocols are provided in respective documentation to provide direction on how to best combine historical Continuous NHANES data with more current iterations.
              </br>
              </br>
              <strong>NOTE:</strong> Due to the coronavirus disease 2019 (COVID-19) pandemic, the 2019-20 Continuous NHANES series was suspended in March 2020. Consequently, only 18 of 30 PSUs were recorded, rendering the sample not nationally representative. To compensate, the 2017-March 2020 Pre-pandemic dataset was created. Users should be aware that combining these programs introduces new challenges due to differences in sampling methods and time spans. For additional details, refer to the [2017–March 2020 Prepandemic File: Sample Design, Estimation, and Analytic Guidelines](https://www.cdc.gov/nchs/data/series/sr_02/sr02-190.pdf) for additional details\ [@Akinbami2017].
            </span>
          </td>
          <td class="var">
            <span class="hidden-content hidden">
              The list below is intended to provide researchers with a sense of the types of variables captured throughout Continuous NHANES. Not all variables were included in each time and some methods differed between bi-annual implementations.
              </br>
              </br>
              <ul>
                <li>Demographic</li>
                <li>Socioeconomic</li>
                <li>Family characteristics</li>
                <li>Sampling weights for those interviewed and examined were adjusted to reflect the participants' respective race, ethnicity, income, and age category in the national population. Note that certain cycles had major changes to their choice of sex-age-race subdomains and oversampling protocols, necessitating additional measures such as weight rescaling when joining different cycles\ [@continuous_analytical_note; @TeChing2020].</li>
                <li>Social support and residence questionnaire for adults</li>
                <li>General physical examination and central nervous system evaluation</li>
                <li>Body measurements</li>
                <li>Dental, vision, and hearing examination</li>
                <li>Respiratory health through exhaled nitric oxide and spirometry</li>
                <li>Blood and urine assessments</li>
                <li>Liver ultrasound transient elastography</li>
                <li>Whole body dual-energy X-ray</li>
                <li>Dietary practices, food frequency and total nutrient intake</li>
                <li>24-Hour recall</li>
                <li>Medications and supplements</li>
                <li>Drug abuse and alcohol use testing and questionnaire</li>
                <li>Tobacco use questionnaire and examination</li>
                <li>Physical activity questionnaire for adults and examination for all participants</li>
                <li>Functional impairment</li>
                <li>Environmental and occupational exposures, such as pesticides, VOCs, and lead</li>
              </ul>
            </span>
          </td>
          <td>
            <span class="hidden-content hidden">
              <ul>
                <li>10,000 to 12,000 were contacted for each two-year cycle. Responses varied with an average 7,000 to 10,500 interviewed and, of these, 7,000 to 10,000 examined\ [@Zipf2013: @Rothwell2014; @Terry2021].</li>
                <li>All ages from birth</li>
                <li>Noninstitutionalized civilian population</li>
                <li>All 50 States</li>
              </ul>
              </br>
              <strong>Oversampling 1999-06:</strong> Non-Hispanic black persons, Mexican-American persons, low-income white persons beginning in 2000 (at or below 130% of the federal poverty line), adolescents aged 12–19, and persons aged 70 years and over. A supplemental sample of pregnant women was also included, but they were not an oversampled group\ [@Zipf2013].
              </br>
              </br>
              <strong>Oversampling 2007-10:</strong> Non-Hispanic black persons, Hispanic persons, low-income persons (at or below 130% of the federal poverty line), persons aged 80 years and over\ [@Zipf2013].
              </br>
              </br>
              <strong>Oversampling 2011-14:</strong> The same groups as the 2007-10 study design with non-Hispanic non-back Asian persons\ [@Rothwell2014].
              </br>
              </br>
              <strong>Oversampling 2015-18:</strong> The same groups as the 2011-14 study design with low-income persons determined at or below 185% of the federal poverty line and persons 0-11 years old\ [@TeChing2020].
              </br>
              </br>
              <strong>Oversampling 2019-20:</strong> The original design included the same groups as the 2011-14 study design without the 0-11 years old group. Due to the premature suspension of operations in March 2020 due to the COVID-19 pandemic, the data are not nationally representative on their own. These data are available as a convenience sample\ [@nhanes_2019_20].
              </br>
              </br>
              <strong>Oversampling 2021-23:</strong> Efforts to limit participant and staff exposure to COVID-19 led to modifications in every aspect of study design and sampling. To reduce the number of dwellings visited, there was no oversampling of racial or socioeconomic groups. Households with any individuals aged 0–19 years and 60 and older were eligible to participate, while not everyone aged 20–59 years from the same household was included\ [@Terry2021].
            </span>
          </td>
        </tr>
  </tbody>
</table>


<script>
document.addEventListener('DOMContentLoaded', function() {
    const buttons = document.querySelectorAll('.toggle-btn');
    buttons.forEach(btn => {
        btn.addEventListener('click', function() {
            const targetRow = this.getAttribute('data-target');
            const cells = document.querySelectorAll(`#${targetRow} .hidden-content`);
            cells.forEach(cell => {
                cell.classList.toggle('hidden');
            });
            this.classList.toggle('toggled');
        });
    });
});
</script>



:::{.callout-caution collapse="true"}
One project plan for the Continuous NHANES was to allow for survey and content updates to meet emerging public health needs. It is challenging to comprehensively summarize each change made since 1999, as the program has been administered every two years since then. A generalized summary with highlights of major changes is provided in the table below. Researchers interested in learning more about detailed protocol changes are encouraged to review the relevant documentation for each year available on the [NHANES Questionnaires, Datasets, and Related Documentation](https://wwwn.cdc.gov/nchs/nhanes/default.aspx) webpage, the brochere linked above, or the [NHANES Survey Methods and Analytic Guidelines](https://wwwn.cdc.gov/nchs/nhanes/analyticguidelines.aspx) webpage, which consolidates some of the program's documentation\ [@data_nhanes; @nhanes_most_protocols].
:::


### Special Focus Programs and Ancillary Studies

Most programs were intended to collect information about the United States population as a whole and its various subgroups, while sometimes oversampling age and income groups expected to be more vulnerable to malnutrition\ [@Rice1973]. The three NHES, often referred to as Cycles I-III, are essentially iterations of NHANES conducted separately to focus on specific age groups individually\ [@nhes_I; @nhes_II; @nhes_III]. Throughout these programs, expert evaluation panels were assembled to assess NHANES's ability to address key issues\ [@Feinleib1985].

In some cases, this warranted the consolidation of existing NHANES and NHES research, the addition of programs to enhance representation, and ancillary studies to address specific scientific inquiries not directly supported by the primary programs. These are summarized below, and additional information can be found on the [NHANES Questionnaires, Datasets, and Related Documentation](https://wwwn.cdc.gov/nchs/nhanes/default.aspx) webpage.


<table class="custom-table">
    <thead>
        <tr>
            <th class="toggle"></th>
            <th class="iteration">Iteration</th>
            <th class="design">Design and Data</th>
            <th class="var">Variable Categories</th>
            <th class="population">Population</th>
        </tr>
    </thead>
    <tbody>
        <!-- First Row -->
        <tr id="row1b">
            <td class="toggle">
              <button class="toggle-btn" data-target="row1b">
                <i class="fa fa-angle-down"></i>
              </button>
            </td>
            <td class="iteration"><a href="https://wwwn.cdc.gov/nchs/data/nhes123/GROWTHCH.htm" target="_blank">Growth Chart (GROWTHCH)</a>\ [@growth_chart]</td>
            <td class="design">
              <span class="hidden-content hidden">
                <strong>Dates:</strong> 1963-94
                </br>
                </br>
                <strong>Study Design:</strong> A multistage, stratified probability sample of loose clusters of individuals in land-based segments was used, with specifics applicable to the year of observation. Participants were selected from NHES II through NHANES III for longitudinal child development observation, intended to construct the 2000 Centers for Disease Control and Prevention (CDC) growth charts.
                </br>
                </br>
                <strong>Data Types:</strong>\ [@Linder1965]
                <ul>
                  <li>Household interview</li>
                  <li>Available health records</li>
                  <li>Direct medical and dental examination (test results and measurements)</li>
                </ul>
                </br>
                <strong>Data Files:</strong> TXT, SAS, and PDF. Data available on the respective NHES or NHANES program's page.
              </span>
            </td>
            <td class="var">
              <span class="hidden-content hidden">
                <ul>
                  <li>NHANES survey number 1 to 5, starting with NHES II</li>
                  <li>Pseudo-primary sampling unit (PSU) (1-210 outcomes)</li>
                  <li>Pseudo-strata (1-49 outcomes)</li>
                  <li>Demographics of the child</li>
                  <li>Adult respondents relationship to child</li>
                  <li>Exam statistical weight</li>
                  <li>Anthropometric measurements</li>
                  <li>Body mass index (kg/m$^2$)</li>
                  <li>Reported and certified birth weight (gm)</li>
                  <li>Reported and certified low and very low birth weight flag</li>
                </ul>
              </span>
            </td>
            <td>
              <span class="hidden-content hidden">
                <ul>
                  <li>0-25 years of age</li>
                  <li>Noninstitutionalized civilian population</li>
                </ul>
                </br>
                <strong>Number of sample persons NHES II (1963-65):</strong> 7,119 persons sampled approximately 6–12 years of age.
              </br>
              </br>
              <strong>NHES III (1966-70):</strong> 6,768 persons sampled approximately 12–18 years of age.
              </br>
              </br>
              <strong>NHANES I (1971-74):</strong> 9,690 persons sampled approximately 1-25 years of age.
              </br>
              </br>
              <strong>NHANES II (1976-80):</strong> 10,102 persons sampled approximately <1-25 years of age.
              </br>
              </br>
              <strong>NHANES III (1988-04):</strong> 16,397 persons sampled approximately <1-25 years of age.
              </span>
            </td>
        </tr>
        <!-- Second Row -->
        <tr id="row2b">
          <td class="toggle">
            <button class="toggle-btn" data-target="row2b"><i class="fa fa-angle-down"></i></button>
          </td>
          <td class="iteration"><a href="https://wwwn.cdc.gov/nchs/nhanes/nhefs/default.aspx" target="_blank">NHANES Epidemiologic Followup Study (NHEFS)</a>\ [@NHEFS]</td>
          <td class="design">
            <span class="hidden-content hidden">
              <strong>Dates:</strong> 1982-84, 1986, 1987, and 1992
              </br>
              </br>
              <strong>Study Design:</strong> These series were designed to investigate the relationships between clinical, nutritional, and behavioral factors assessed in NHANES I and subsequent morbidity, mortality, and hospital utilization, as well as changes in risk factors, functional limitation, and institutionalization. 90-94% qualifying participants from NHANES I were traced for follow-up\ [@NHEFS].
              </br>
              </br>
              <strong>Data Types:</strong>
                <ul>
                  <li>Personal or computer-assisted telephone interview with subjects or their proxies</li>
                  <li>Physical examination for 1982-84 only</li>
                  <li>Hospital and nursing home records of overnight stays</li>
                  <li>Death certificates of decedents</li>
                </ul>
              </br>
              <strong>Data Files:</strong> TXT, SAS, and PDF
            </span>
          </td>
          <td class="var">
            <span class="hidden-content hidden">
              <ul>
                <li>Demographic (including twin status)</li>
                <li>Socioeconomic</li>
                <li>Family characteristics</li>
                <li>Geographic (state and urbanicity)</li>
                <li>Sampling weights to reflect 24 million children of the base population</li>
                <li>Medical and developmental history</li>
                <li>School questionnaire (i.e. academic standing, use of special resources)</li>
                <li>Psychological examination</li>
                <li>Physical examination</li>
                <li>Ear, nose, and throat examination</li>
                <li>Diagnostic impressions (10 records including neurological)</li>
                <li>Audiometry</li>
                <li>Vision</li>
                <li>Dental findings</li>
                <li>Body measurements (13 sitting and 13 standing)</li>
                <li>Bone age by X-ray</li>
              </ul>
            <span class="hidden-content hidden">
          </td>
          <td>
            <span class="hidden-content hidden">
              <ul>
                <li>14,407 persons from NHANES I who were examined</li>
                <li>25-74 years of age</li>
                <li>Noninstitutionalized civilian population</li>
                <li>Sampling was restricted to the conterminous United States and excluded individuals residing in Native American reservations.</li>
                <li>Emphasis by oversampling of the following groups: at or below the federal poverty line, preschool children (6 months - 5 years), women of child-bearing age, and persons aged 60-74 years old.</li>
                </br>
                </br>
                <strong>1986 NHEFS</strong>
                </br>
                </br>
                <strong></strong>
                </br>
                </br>
                <strong></strong>
                The 1986 NHEFS was conducted for members of the NHEFS cohort who were 55-74 years of age at their baseline examination and not known to be deceased at the 1982-1984 NHEFS (n = 3,980). The 1987 NHEFS was conducted for the entire nondeceased NHEFS cohort (n = 11,750). The fourth wave of data collection, the 1992 NHEFS, includes the entire nondeceased NHEFS cohort (n = 11,195).
              </ul>
            </span>
          </td>
        </tr>
        <!-- Third Row -->
        <tr id="row3b">
          <td class="toggle">
            <button class="toggle-btn" data-target="row3b"><i class="fa fa-angle-down"></i></button>
          </td>
          <td class="iteration"><a href="https://wwwn.cdc.gov/nchs/nhanes/search/nnyfs12.aspx" target="_blank">NHANES National Youth Fitness Survey (NNYFS)</a>\ [@nhes_III]</td>
          <td class="design">
            <span class="hidden-content hidden">
              <strong>Dates:</strong> 2012
              </br>
              </br>
              The study design, data types, and available files types are the same as those of NHES II\ [@Woolsey1974].
            </span>
          </td>
          <td>
            <span class="hidden-content hidden">
              Variables tracked are the same as NHES II.
            </span>
          </td>
          <td>
            <span class="hidden-content hidden">
              <ul>
                <li>7,514 interviewed with 6,768 examined</li>
                <li>12-17 years of age</li>
                <li>Parents from the noninstitutionalized civilian population</li>
                <li>Sampling was restricted to the conterminous United States.</li>
              </ul>
              </br>
              <strong>NOTE:</strong> Nearly 1/3$^{rd}$ of the children examined in NHES II were reexamined in NHES III. The Extended Data Tape (EDT) helps the user identify the entries present in both cycles to create a composite data set for longitudinal studies.
            </span>
          </td>
        </tr>
        <!-- Fourth Row -->
        <tr id="row4b">
          <td class="toggle">
            <button class="toggle-btn" data-target="row4b"><i class="fa fa-angle-down"></i></button>
          </td>
          <td class="iteration"><a href="https://wwwn.cdc.gov/nchs/nhanes/nhanes3/default.aspx" target="_blank">NHANES III Augmented Data, Special Studies, and Surplus Sera Data</a>\ [@nhanes_III]</td>
          <td class="design">
            <span class="hidden-content hidden">
              <strong>Dates:</strong> Multiple follow-ups after completion of NHANES III in 1994.
              </br>
              </br>
              <strong>Study Design:</strong> Samples of persons were selected through complex, multistage, highly stratified, cluster samples of households. Similarly to NHES, participants were randomly selected from primary sampling units (PSUs) that were partitioned into segments based on geography, eligibility, and housing density\ [@Linder1965].
              </br>
              </br>
              <strong>Data Types:</strong>\ [@Harvey1981]
              </br>
              The following was collected from all participants
              <ul>
                <li>Household interview</li>
                <li>Available health records and history</li>
                <li>Direct medical, dental, and dermatological examination (test results and measurements)</li>
                <li>Vision and hearing tests</li>
                <li>Anthropometric measurements</li>
                <li>Hand-wrist X-rays for participants ages 1-17</li>
                <li>Nutrient composition, frequency, and adequacy</li>
              </ul>
              </br>
              Detailed examinations included the above list and
              <ul>
                <li>Supplement medical history for arthritis, respiratory and cardiovascular conditions</li>
                <li>Health care needs questionnaire</li>
                <li>General well-being questionnaire</li>
                <li>Extended examination and tests</li>
                <li>X-rays of the chest and hip and knee joints</li>
              </ul>
              </br>
              <strong>Data Files:</strong> TXT, SAS, or XPT and PDF
            </span>
          </td>
          <td class="var">
            <span class="hidden-content hidden">
              Additional details concerning the two baseline populations are summarized in the [_Plan and Operation of the Health and Nutrition Examination Survey Series 1 Number 10a_](https://www.cdc.gov/nchs/data/series/sr_01/sr01_010a.pdf) PDF pg. 20.
              </br>
              </br>
              <ul>
                <li>Demographic</li>
                <li>Socioeconomic</li>
                <li>Family characteristics</li>
                <li>Geographic (state and urbanicity)</li>
                <li>Sampling weights to reflect the population of 1-74 year olds within 60 age-sex-race groups at the midpoint of the survey</li>
                <li>Medical history questionnaire, ages 1-11 years,</li>
                <li>Medical history questionnaire, ages 12-74 years</li>
                <li>Medical examination, ages 1-74 years</li>
                <li>Health care needs, general medical history supplement, and respiratory and cardiovascular supplements, ages 25-74 years</li>
                <li>General well-being and the CES-D depression scale developed by The National Institute of Mental Health, ages 25-74 years</li>
                <li>Dental, ages 1-74 years</li>
                <li>Dermatology, ages 1-74 years</li>
                <li>Near and distant vision, ages 25-74 years</li>
                <li>Ophthalmology, ages 1-74 years</li>
                <li>Audiometric</li>
                <li>Spirometry-best trials only, ages 25-74 years</li>
                <li>Anthropometry goniometry, skeletal age, bone density, and cortical thickness, Ages 1-74 years</li>
                <li>Arthritis</li>
                <li>Chest X-ray, pulmonary diffusion, and tuberculin test results, ages 25-74 years</li>
                <li>Electrocardiogram</li>
                <li>Computer measurements and interpretations of electrocardiograms, ages 25-74 years</li>
                <li>Biochemistry, serology, hematology, blood slides, urine dipstick</li>
                <li>Model gram</li>
                <li>Nutrient composition</li>
                <li>Dietary frequency and adequacy, ages 1-74 years</li>
                <li>24-hour food consumption intake, ages 1-74 years</li>
              </ul>
            </span>
          </td>
          <td>
            <span class="hidden-content hidden">
              <ul>
                <li>32,000 were contacted with 31,973 interviewed and, of these, 23,808 examined</li>
                <li>Noninstitutionalized civilian population</li>
                <li>Sampling was restricted to the conterminous United States and excluded individuals residing in Native American reservations.</li>
              </ul>
              </br>
              The base population surveyed and examined was divided into two groups of interest: one primarily focused on nutrition, and a subset of this group that underwent a more detailed process. Each of these had different sampling emphases, which is summarized below\ [@Rice1973; @nhanes_I].
              </br>
              </br>
              <strong>Standard examination cohort:</strong>
              </br>
              <ul>
                <li>1-74 years of age</li>
                <li>Emphasis by oversampling of the following groups: at or below the federal poverty line, preschool children (6 months - 5 years), women of child-bearing age, and persons aged 60-74 years old.</li>
              </ul>
              </br>
              <strong>Detailed examination cohort:</strong>
              </br>
              <ul>
                <li>6,913 of the 32,000. 3,059 received detailed medical examinations during the survey and an additional 3,059 received detailed examinations after completion of the nutrition survey in October 1975. The last group is referred to as the NHANES I Augmentation group.</li>
                <li>25-74 years of age</li>
                <li>No oversampling relative to the national population</li>
              </ul>
            </span>
          </td>
        </tr>
  </tbody>
</table>


#### Hispanic Health and Nutrition Examination Survey (HHANES) 1982-84

After the completion of NHANES II in 1980, an expert panel recommended that the next program focus on the health and nutritional status of a specific subpopulation. Different groups were considered, and it was determined that the Hispanic population should be the focus due to notable gaps in comparable data for this population. This was partly due to insufficient representation of Hispanic populations and a lack of separate representation for Mexican-American, Cuban-American, and Puerto Rican communities. HHANES was formulated with the explicit purpose of supporting estimations of the health and nutritional status of the United States Hispanic population so that it is comparable with the general population, while also considering variations between these three specific Hispanic communities\ [@Feinleib1985].

Additional study design and content details are provided earlier in the [Methods and Representational Differences](#methods-and-representational-differences) section.

::: {.callout-note collapse="true"}
The chapter dedicated to cross-cultural considerations, found in the [Plan and Operation of HHANES](https://www.cdc.gov/nchs/data/series/sr_01/sr01_019.pdf) documentation, acknowledged a need for cultural sensitivity and recognized unknown obstacles to implementation. It specifically noted that insufficient personnel on the core team were from or otherwise knowledgeable about the Hispanic population and cultures. This chapter highlighted that this was the first known attempt by the National Center for Health Statistics (NCHS) to conduct a survey focused on a specific ethnic group. This document details the strategies the team employed to address these shortcomings by hiring qualified personnel and forming task force committees intended to closely evaluate aspects of the program, from translation to survey content\ [@Feinleib1985].
:::


## Gaining Access

Freely accessible vs restricted access

### Do I Qualify?

### Step-by-Step

### Typical Timeline


::: {.panel-tabset}

## Freely Accessible

Links and time

## Restrcited Access

Links and time

## Biospecimens

Links and time

:::

## Publications

```{r}
#| echo: false

# Fetch and format the citations.
citations <- get_citations(
  keywords = c("National Health and Nutrition Examination Survey", "NHANES"), 
  affiliation = "Yale School of Public Health")

# Condition the output based on presence of articles.
citations_html <- make_pub_list(citations)

if(!any(str_detect(citations$citations, "No publications"))) {
  citation_length <- length(citations$citations)
  citation_date   <- str_match(citations$query, "\\((\\d{4}):(\\d{4})\\[dp\\]\\)")[,2]
} else {
  citation_length <- 0
  citation_date   <- as.numeric(format(Sys.Date(), "%Y")) - as.numeric(str_extract("No publications found in the last 25 years.", "[0-9]{2}"))
}

```


This section displays PubMed articles that mention the dataset and are authored by individuals affiliated with the Yale School of Public Health. Please note that the PubMed API does not distinguish between uppercase and lowercase characters when processing queries. As a result, there may be occasional mismatches in the search results.

`r citation_length` citations released after `r citation_date` were found.

::: {style="margin-left: 1rem"}

```{r}
#| echo: false
#| results: 'asis'

# Display formatted citations.
cat(citations_html)
```

```{=html}
<!-- Pagination buttons and scripting -->
<div class="citations-navigation">
  <button onclick="prevPage()">&#9664; Previous</button>
  <button onclick="nextPage()">Next &#9654;</button>
</div>

<script>
  document.addEventListener("DOMContentLoaded", function() {
    let currentPage = 1;
    const itemsPerPage = 3;
    const citationsContent = document.getElementById("citations-content");

    if (citationsContent) {
      const numPages = Math.ceil(citationsContent.children.length / itemsPerPage);

      function updateCitations() {
        const items = citationsContent.children;
        for (let i = 0; i < items.length; i++) {
          items[i].style.display = (i >= (currentPage - 1) * itemsPerPage && i < currentPage * itemsPerPage) ? "list-item" : "none";
        }
      }

      function nextPage() {
        if (currentPage < numPages) {
          currentPage++;
          updateCitations();
        }
      }

      function prevPage() {
        if (currentPage > 1) {
          currentPage--;
          updateCitations();
        }
      }

      // Initial call to update citations on load
      updateCitations();

      // Attach functions to window object for button actions
      window.nextPage = nextPage;
      window.prevPage = prevPage;
    }
  });
</script>
```

::: {.callout-note collapse="true"}
The search PubMed query used to generate this list is:

`r citations$query`
:::

:::