---
  title: "NHANES"
  description: |
         </br>
         The National Health and Nutrition Examination Survey (NHANES) collects data about the health of adults and children in the United States. We also collect data about what participants eat, drink, and take as supplements to determine how many nutrients are in their diet. These dietary interviews and blood tests help us measure the nutritional status of U.S. adults and children.
         </br>
         </br>
         CDC's National Center for Health Statistics (NCHS) conducts NHANES. NHANES is the only national health survey that includes health exams, laboratory tests, and dietary interviews for participants of all ages.
         </br>
         </br>
         NHANES data can help improve the health of Americans. Survey data have driven changes in how doctors treat patients and how public policy supports good health.
  title-block-banner: true
  format: html
  filters:
    - parse-latex
  markdown: 
  editor_options: 
  chunk_output_type: console
---

```{r}
#| echo: false

# Import necessary libraries
suppressPackageStartupMessages({
  library("stringr")
  library("htmltools")
  library("glue")
  library("rentrez")
  library("xml2")
  library("clipr")
})

"%!in%" <- function(x,y)!("%in%"(x,y))

# Load in the functions.
source('./PubMed_Search_Functions.R')
```


## Publications

```{r}
#| echo: false

# Fetch and format the citations.
citations <- get_citations(
  keywords = c("National Health and Nutrition Examination Survey", "NHANES"), 
  affiliation = "Yale School of Public Health")

# Condition the output based on presence of articles.
citations_html <- make_pub_list(citations)
if(!any(str_detect(citations$citations, "No publications"))) {
  citation_length <- length(citations$citations)
  citation_date   <- str_match(citations$query, "\\((\\d{4}):(\\d{4})\\[dp\\]\\)")[,2]
} else {
  citation_length <- 0
  citation_date   <- as.numeric(format(Sys.Date(), "%Y")) - as.numeric(str_extract("No publications found in the last 25 years.", "[0-9]{2}"))
}

```


This section displays PubMed articles that mention the dataset and are authored by individuals affiliated with the Yale School of Public Health. Please note that the PubMed API does not distinguish between uppercase and lowercase characters when processing queries. As a result, there may be occasional mismatches in the search results.

`r citation_length` citations released after `r citation_date` were found.

::: {style="margin-left: 1rem"}

```{r}
#| echo: false
#| results: 'asis'

# Display formatted citations.
cat(citations_html)
```

```{=html}
<!-- Pagination buttons and scripting -->
<div class="citations-navigation">
  <button onclick="prevPage()">&#9664; Previous</button>
  <button onclick="nextPage()">Next &#9654;</button>
</div>

<script>
  document.addEventListener("DOMContentLoaded", function() {
    let currentPage = 1;
    const itemsPerPage = 3;
    const citationsContent = document.getElementById("citations-content");

    if (citationsContent) {
      const numPages = Math.ceil(citationsContent.children.length / itemsPerPage);

      function updateCitations() {
        const items = citationsContent.children;
        for (let i = 0; i < items.length; i++) {
          items[i].style.display = (i >= (currentPage - 1) * itemsPerPage && i < currentPage * itemsPerPage) ? "list-item" : "none";
        }
      }

      function nextPage() {
        if (currentPage < numPages) {
          currentPage++;
          updateCitations();
        }
      }

      function prevPage() {
        if (currentPage > 1) {
          currentPage--;
          updateCitations();
        }
      }

      // Initial call to update citations on load
      updateCitations();

      // Attach functions to window object for button actions
      window.nextPage = nextPage;
      window.prevPage = prevPage;
    }
  });
</script>
```

::: {.callout-note collapse="true"}
The search PubMed query used to generate this list is:

`r citations$query`
:::

:::

\
\
\
Updated: June 15th, 2025