---
  title: |
         Epic Cosmos
  description: |
         </br>
         The Cosmos data set combines billions of clinical data points, made up of data from across the Epic community, to form the largest database of Electronic Health Record (EHR) patient information. Participating organizations provide a HIPAA-defined limited data set to the centralized Cosmos database. 
         </br>
         </br>
         With over 34,200 Clinics and 1,522 Hospitals participating in Cosmos, researchers can access a representative sample of patients who seek healthcare. You can learn more about the latest features of Cosmos on <a href="https://signin.epic.com/adfs/ls/?wa=wsignin1.0&wtrealm=https%3a%2f%2fuserweb.epic.com%2f&wctx=rm%3d1%26id%3dpassive%26ru%3d%252F&wct=2025-04-16T00%3a32%3a32Z&wreply=https%3a%2f%2fuserweb.epic.com%2f">Epic UserWeb</a>, available to all Yale New Haven Health System (YNHHS) research users and on the <a href="https://teams.microsoft.com/l/team/19%3AJTlr92eAMEnJsBJmV1Uymbla8oek8ePqASKzxPE7mrQ1%40thread.tacv2/conversations?groupId=aa0ac073-1651-4dee-a59e-884bef65ea6b&tenantId=dd8cbebb-2139-4df8-b411-4e3e87abeb5c">Yale Cosmos Teams</a> Site.
  title-block-banner: true
  format: html
  editor: source
  filters:
    - parse-latex
  markdown: 
  editor_options: 
  chunk_output_type: console
---

```{r}
#| echo: false

# Import necessary libraries
suppressPackageStartupMessages({
  library("stringr")
  library("htmltools")
  library("glue")
  library("rentrez")
  library("xml2")
  library("clipr")
})

"%!in%" <- function(x,y)!("%in%"(x,y))

# Load in the functions.
source('./PubMed Search_Functions.R')
```

<small>Updated: November 5<sup>th</sup>, 2025</small>


the proportions of patients represented in Cosmos compare well to the United States Census.


There are three levels of access available:

::: {style="margin-left: 2rem"}
-   RBA without an IRB give access to the afore mentioned resources but only supplies de-identified, summary count data produced within YNHHS Epic.
-   RBA with an approved IRB gives access to individual-level data for Epic data produced within YNHHS that is strictly relevant to the IRB approved study.
-   Epic Cosmos - RBA holders can gain access to de-identified, summary count data produced by Epic participating institutions outside of YNHHS upon special request.
:::




## Gaining Access

::: {.panel-tabset}

### Researcher Basic Access (RAB)

#### Do I Qualify?

Yale University users with Research Basic Access (RBA) can apply for access to Epic Cosmos data.

#### Typical Timeline

It takes a couple of days for CITI trainings to show up on Workday. Completed ceritifcates take no more than 2 business days to be reflected on Workday. Time for recognition of completion varies if not done through Workday.

Note that CITI training is involved and may take between 1-2 weeks to complete.


#### Step-by-Step Guide

If you haven't obtained RBA yet, please follow the instructions provided on the [YNHHS Epic](YNHHS-Epic.html) webpage.

**a: Complete Yale's HIPAA Training**

**For Yale NetID holders:** Complete the [HIPAA Privacy and Security](https://hipaa.yale.edu/training/training-modules) courses in Workday, including the basic privacy and security module followed by the role-specific module (e.g., for clinicians or researchers).

**For Non-Yale NetID holders:** Read and sign the [HIPAA Privacy and Security](https://hipaa.yale.edu/training/training-modules) training materials specific to your role on the project (e.g., for clinicians or researchers). Submission directions for these documents can be found on the PDF.


**b: Obtain a Human Subjects Protection Training (HSPT) Certificate from CITI**

Human Subjects Protection Training (HSPT) certificate from CITI that has been completed within the last 3 years. Below is a summary of the process, and the complete directions can be found on the [It's Your Yale](https://your.yale.edu/research-support/human-research-protection-program/education-and-training/human-research-training) page about HSPT.

Researchers will need to complete the following CITI certificates:

::: {style="margin-left: 2rem"}
-   "Basic Training"
-   One of the following two initial/basic Human Subjects Research modules: "Biomedical Research Basic" or "Social Behavioral Educational Researchers Basic".
:::

If you have already completed your CITI certification through another institution, please send a copy to [irb.training@yale.edu](mailto:irb.training@yale.edu) and skip to Section C. For further questions, you can also contact this email address.

**For Yale Employee:** It is preferable, though not required, to access the CITI training modules through Workday. This will expedite the recognition of the certificate in the Yale Integrated Research Enterprise Solution (IRES) IRB system.

::: {style="margin-left: 2rem"}
1. Open the [Learning Homepage](https://www.myworkday.com/yale/learning) in Workday.

2. The CITI training link will appear as the "Human Subjects Protection Training – Initial/Refresher" course under the Required Training section. If it does not appear, i) ensure your name is listed in the IRES IRB pending study or submission, ii) reach out to your manager, or iii) contact [learning@yale.edu](mainto:learning@yale.edu) or [irb.training@yale.edu](mainto:irb.training@yale.edu) to request a manual assignment.

:::{.callout-note}
If available, you may need to complete the "Compliance Assessment" to indicate that you are conducting human subjects research.
:::

3. Once you access the "Human Subjects Protection Training – Initial/Refresher" course page, click the "Enroll" button to proceed. This will change the course status to **Start Program**.

4. A button labeled **Start Program**, or **Retake Program** if you are taking this as a refresher, will appear under the course photo. Click it to proceed.

5. Log into [CITI](https://www.citiprogram.org/index.cfm?pageID=668) using the "Yale University Training Opportunities" Single Sign-On (SSO).
:::

**For Non-Yale Employee:** If you are not a Yale employee, you will need to access CITI through your institution or log in directly from their [main webpage](https://about.citiprogram.org/). Please be aware that pursuing certification independently may incur charges. You can find the current costs for CITI courses on their [Explore Our Courses](https://about.citiprogram.org/courses/?reset=true) webpage.

**For All Applicants:** 

The previous steps guide applicants on how to access CITI courses and detail how Yale employees can expedite recognition of their certification by using Workday. Once you have access to a CITI account, follow these steps to complete your certification.

::: {style="margin-left: 2rem"}
a. In CITI, click “Add a Course” and scroll to the bottom to find the initial/basic CITI HSP Training under the Human Subjects Research heading. 

b. Indicate your intent to take a Human Subjects Research course and select either “Biomedical Research” or “Social Behavioral Educational Researchers”. When prompted, confirm that you want to receive the basic course, specifically the "Biomedical Research-Basic Course".

c. Complete the course modules and attain a passing grade of at least 70% on the assessment.

d. If you did not link your Workday account with CITI, save the completion certificate and send it to [irb.training@yale.edu](mainto:irb.training@yale.edu). If you accessed CITI through Workday, your record should auto-update.
:::


**c: **

:::


:::


## Valuable Links

::: {style="margin-left: 2rem"}
- [YNHHS Learning Management System (LMS)](https://ynhh.certpointsystems.com/): Center Point is where you can sign up for in-person or online training courses.

- [MyAPPs](https://myapps.ynhh.org): Find the portal for accessing the VDI using Citrix Workspace. Ensure that YNHHS DUO and YU/YNHHS VPN are utilized if access is attempted outside the Yale University or YNHHS network.

- [JDAT Analytics Portal](https://jdat.ynhh.org/AnalyticsPortal): Submit requests to the JDAT team, make Helix requests (Yale’s customized data warehouse system), and find YNHHS-created training for tools like Epic's Slicer Dicer.

- [Epic UserWeb](https://userweb.epic.com/): Find Epic training materials, documentation, community discussions, support resources, updates, customization options, and access to your Epic profile.

- [Epic University](https://training.epic.com/): A subdomain of the UserWeb, find training opportunities and access to coursework necessary to obtain certificates, such as the Super User Badge, with courses organized by the Epic team.

- [Galaxy](https://galaxy.epic.com/?#Home): A subdomain of the UserWeb, find comprehensive resource that provides documentation, announcements, best practices, and implementation guides related to the Epic dataset.

:::


## Important Links

::: {style="margin-left: 2rem"}
- [YNHHS Learning Management System (LMS)](https://ynhh.certpointsystems.com/): Center Point is where you can sign up for in-person or online training courses.
- [MyAPPs](https://myapps.ynhh.org): Access to the VDI via a Citrix Workspace download. YNHHS DUO and YU/YNHHS VPN is required if not accessing on the YU or YNHHS network.

:::

## Publications

```{r}
#| echo: false

# Fetch and format the citations.
citations <- get_citations(
  doi = c("10.7326/M21-0065"),
  affiliation = "Yale School of Public Health"
)

# Condition the output based on presence of articles.
citations_html <- make_pub_list(citations)

if(!any(str_detect(citations$citations, "No publications"))) {
  citation_length <- length(citations$citations)
  citation_date   <- str_match(citations$query, "\\((\\d{4}):(\\d{4})\\[dp\\]\\)")[,2]
} else {
  citation_length <- 0
  citation_date   <- as.numeric(format(Sys.Date(), "%Y")) - as.numeric(str_extract("No publications found in the last 25 years.", "[0-9]{2}"))
}

```


This section presents a selection of PubMed articles that utilize the dataset and are authored by individuals affiliated with the Yale School of Public Health. These articles are provided to inspire researchers and students to use the data in their own work.

::: {style="margin-left: 1rem"}

```{r}
#| echo: false
#| results: 'asis'

# Display formatted citations.
cat(citations_html)
```

```{=html}
<!-- Pagination buttons and scripting -->
<div class="citations-navigation">
  <button onclick="prevPage()">&#9664; Previous</button>
  <button onclick="nextPage()">Next &#9654;</button>
</div>

<script src="../../../Code/pagination.js"></script>
```

:::


## References

The content provided here is a compilation of work from multiple groups throughout YU and YNHHS:

::: {style="margin-left: 2rem"}
1.  [Yale Biomedical Informatics & Computing](https://medicine.yale.edu/ybic)

:::
