---
title-block-banner: true
title: |
  Epic Cosmos
description: |
  </br>
  The Cosmos data set combines billions of clinical data points, made up of data from across the Epic community, to form the largest database of Electronic Health Record (EHR) patient information. Participating organizations provide a HIPAA-defined limited data set to the centralized Cosmos database. 
  </br>
  </br>
  With over 34,200 Clinics and 1,522 Hospitals participating in Cosmos, researchers can access a representative sample of patients who seek healthcare. You can learn more about the latest features of Cosmos on <a href="https://signin.epic.com/adfs/ls/?wa=wsignin1.0&wtrealm=https%3a%2f%2fuserweb.epic.com%2f&wctx=rm%3d1%26id%3dpassive%26ru%3d%252F&wct=2025-04-16T00%3a32%3a32Z&wreply=https%3a%2f%2fuserweb.epic.com%2f">Epic UserWeb</a>, available to all Yale New Haven Health System (YNHHS) research users and on the <a href="https://teams.microsoft.com/l/team/19%3AJTlr92eAMEnJsBJmV1Uymbla8oek8ePqASKzxPE7mrQ1%40thread.tacv2/conversations?groupId=aa0ac073-1651-4dee-a59e-884bef65ea6b&tenantId=dd8cbebb-2139-4df8-b411-4e3e87abeb5c">Yale Cosmos Teams</a> Site.
include-in-header: ../../../_includes/feedback_popup.html
bibliography: BibTeX/ynhhs_epic_references.bib
csl: ../../../Code/nature.csl
---

```{r}
#| echo: false

# Load in the functions.
source('../../../Code/PubMed Search_Functions.R')
```

<small>Updated: November 5<sup>th</sup>, 2025</small>


the proportions of patients represented in Cosmos compare well to the United States Census.


There are three levels of access available:

::: {style="margin-left: 2rem"}
-   RBA without an IRB give access to the afore mentioned resources but only supplies de-identified, summary count data produced within YNHHS Epic.
-   RBA with an approved IRB gives access to individual-level data for Epic data produced within YNHHS that is strictly relevant to the IRB approved study.
-   Epic Cosmos - RBA holders can gain access to de-identified, summary count data produced by Epic participating institutions outside of YNHHS upon special request.
:::

## Overview

The Epic Cosmos data combines inpatient and outpatient charts into a single, longitudinal patient record, removing any duplicated records that may arise from being seen at multiple Cosmos organizations. The data provided to researchers adheres to [HIPAA-defined limited data set](https://privacyruleandresearch.nih.gov/pr_08.asp) rules. In Cosmos, the de-identified dataset enhances patient privacy protections by suppressing geographic information, consistently shifting patient dates, and concealing the source organization. More details can be found in their Data Dictionary, which provides notes on the transformation of individual data elements.

:::{.callout-tip}
## Leverage Expert Analysis for Your Research Needs!

The Joint Data Analytics Team (JDAT) offers customized reporting and data analysis from the Epic data system, specializing in research support. With a team of 60 informaticists and analysts, JDAT works closely with users to understand their data needs and efficiently retrieve information. JDAT supports projects such as cohort population counts, grant application statistics, retrospective reviews, custom data extracts, and report development\ [@ycci_epic_ehr].

Click [here](https://medicine.yale.edu/ybic/data-resources-services/how-to-submit-a-research-data-request/) for directions on submitting a Helix research data request with JDAT, Yale’s customized data warehouse system. Learn more on the [Yale Center for Clinical Investigation](https://medicine.yale.edu/center-clinical-investigation/)'s [Epic Electronic Health Record](https://yaleedu.sharepoint.com/sites/YaleCenterClinicalInvestigation/SitePages/Epic-(EHR).aspx) SharePoint page, which is also accessible from their [For Researchers & Research Professionals](https://medicine.yale.edu/center-clinical-investigation/for-researchers-and-research-staff/) webpage\ [@jdat_helix; @ycci_epic_ehr; @ybic_for_researchers].
:::

The types of data collected in the EMR are continuously expanding. For example, researchers currently have access to the following data\ [@epic_overview]:

::: {style="margin-left: 2rem"}
- **Patient Information:** Limited demographic details such as age, sex, gender identity, race, ethnicity, location, patient status, and family history, social history elements like tobacco and substance use, birth control, social determinants of health, allergy information, and detailed birth history.

- **Visits and Encounters:** Patient encounters and admissions including encounter type, classification, department and provider specialties, and visit reasons, as well as clinical notes from various medical documentation and patient experience data from MyChart usage and patient-reported outcomes.

- **Medical Data:** Diagnoses, clinical and billed procedures, and vital signs such as blood pressure, pulse, temperature, respiratory rate, SpO2 levels, height, weight, BMI, and head circumference.

- **Lab and Genomics:** Lab results such as component-specific results, susceptibility results, and reference ranges, in addition to genomic variant data like allelic state and chromosome details, and biosamples for genomic assays including blood, saliva, and urine samples.

- **Health Assessments:** Information on cancer staging, immunizations, and eye exams, with details on stage group, immunization type, and visual acuity.

- **Specialty Data:** Records data on hospital-acquired infections and various infection types, surgical procedures, transplant data for liver and kidney transplants, and health tracking data from wearable devices.

- **Research and Analytics:** Geographic factors like RUCA and Social Vulnerability Index, risk scores for conditions such as cardiovascular disease and diabetes, and data from research studies detailing study type, enrollment, and partnered research contributions.

- **Miscellaneous:** Encompasses peer group information on various types of hospitals, medication data including prescriptions and patient-reported medications, and additional lab results.
:::


## Gaining Access

::: {.panel-tabset}

### Researcher Basic Access (RAB)

#### Do I Qualify?

Yale University users with Research Basic Access (RBA) can apply for access to Epic Cosmos data.

#### Typical Timeline

It takes a couple of days for CITI trainings to show up on Workday. Completed ceritifcates take no more than 2 business days to be reflected on Workday. Time for recognition of completion varies if not done through Workday.

Note that CITI training is involved and may take between 1-2 weeks to complete.


#### Step-by-Step Guide

If you haven't obtained RBA yet, please follow the instructions provided on the [YNHHS Epic](YNHHS-Epic.html) webpage.

**a: Complete Yale's HIPAA Training**

**For Yale NetID holders:** Complete the [HIPAA Privacy and Security](https://hipaa.yale.edu/training/training-modules) courses in Workday, including the basic privacy and security module followed by the role-specific module (e.g., for clinicians or researchers).

**For Non-Yale NetID holders:** Read and sign the [HIPAA Privacy and Security](https://hipaa.yale.edu/training/training-modules) training materials specific to your role on the project (e.g., for clinicians or researchers). Submission directions for these documents can be found on the PDF.


**b: Obtain a Human Subjects Protection Training (HSPT) Certificate from CITI**

Human Subjects Protection Training (HSPT) certificate from CITI that has been completed within the last 3 years. Below is a summary of the process, and the complete directions can be found on the [It's Your Yale](https://your.yale.edu/research-support/human-research-protection-program/education-and-training/human-research-training) page about HSPT.

Researchers will need to complete the following CITI certificates:

::: {style="margin-left: 2rem"}
-   "Basic Training"
-   One of the following two initial/basic Human Subjects Research modules: "Biomedical Research Basic" or "Social Behavioral Educational Researchers Basic".
:::

If you have already completed your CITI certification through another institution, please send a copy to [irb.training@yale.edu](mailto:irb.training@yale.edu) and skip to Section C. For further questions, you can also contact this email address.

**For Yale Employee:** It is preferable, though not required, to access the CITI training modules through Workday. This will expedite the recognition of the certificate in the Yale Integrated Research Enterprise Solution (IRES) IRB system.

::: {style="margin-left: 2rem"}
1. Open the [Learning Homepage](https://www.myworkday.com/yale/learning) in Workday.

2. The CITI training link will appear as the "Human Subjects Protection Training – Initial/Refresher" course under the Required Training section. If it does not appear, i) ensure your name is listed in the IRES IRB pending study or submission, ii) reach out to your manager, or iii) contact [learning@yale.edu](mainto:learning@yale.edu) or [irb.training@yale.edu](mainto:irb.training@yale.edu) to request a manual assignment.

:::{.callout-note}
If available, you may need to complete the "Compliance Assessment" to indicate that you are conducting human subjects research.
:::

3. Once you access the "Human Subjects Protection Training – Initial/Refresher" course page, click the "Enroll" button to proceed. This will change the course status to **Start Program**.

4. A button labeled **Start Program**, or **Retake Program** if you are taking this as a refresher, will appear under the course photo. Click it to proceed.

5. Log into [CITI](https://www.citiprogram.org/index.cfm?pageID=668) using the "Yale University Training Opportunities" Single Sign-On (SSO).
:::

**For Non-Yale Employee:** If you are not a Yale employee, you will need to access CITI through your institution or log in directly from their [main webpage](https://about.citiprogram.org/). Please be aware that pursuing certification independently may incur charges. You can find the current costs for CITI courses on their [Explore Our Courses](https://about.citiprogram.org/courses/?reset=true) webpage.

**For All Applicants:** 

The previous steps guide applicants on how to access CITI courses and detail how Yale employees can expedite recognition of their certification by using Workday. Once you have access to a CITI account, follow these steps to complete your certification.

::: {style="margin-left: 2rem"}
a. In CITI, click “Add a Course” and scroll to the bottom to find the initial/basic CITI HSP Training under the Human Subjects Research heading. 

b. Indicate your intent to take a Human Subjects Research course and select either “Biomedical Research” or “Social Behavioral Educational Researchers”. When prompted, confirm that you want to receive the basic course, specifically the "Biomedical Research-Basic Course".

c. Complete the course modules and attain a passing grade of at least 70% on the assessment.

d. If you did not link your Workday account with CITI, save the completion certificate and send it to [irb.training@yale.edu](mainto:irb.training@yale.edu). If you accessed CITI through Workday, your record should auto-update.
:::

**c: Accessing Epic Cosmos**


Accessing the Epic Cosmos SlicerDicer is done through the Epic Cosmos portal, and not the YNHH VDI. You can either open this portal from the Epic UserWeb by navigating to the "Show All" drop down menu and selecting "Cosmos" under "Support". Or, you can directly access Cosmos at [cosmos.epichosted.com](https://cosmos.epichosted.com/EpicCareLink_AUTH/common/epic_check.asp).


**d: Training**

Training based on role. Can list out the ones for "Generating reports" and "research"

"Cosmos Portal & Analytics Resources" in Galaxy for additional learning materials for using Epic and it's analytical tools like SlicerDicer

[Library of Learning Dashboards](https://cosmos.epichosted.com/EpicCareLink_AUTH/common/epic_check.asp)

#### Certifications


#### Publications Expectations

Epic provides data users with a checklist for preparing data for publications, emphasized in the user agreement, "Cosmos Rules of the Road". The full instructions and expectations can be found in the Galaxy "Cosmos Publication Checklist" article and the "Cosmos Rules of the Road", which outlines some of the Data Use Agreement terms\ [@epic_publication_checklist].

::: {.callout-tip}
## Data Privacy: Low Count Suppression

When transferring results off the VDI or sharing SlicerDicer graphs, it is imperative to follow Epic's process to prevent the exportation of line-level data. No counts less than 10 can be individually noted (report as <11).

Review the [Centers for Medicare and Medicaid Services (CMS) Cell Size Suppression Policy](https://resdac.org/articles/cms-cell-size-suppression-policy) outlined by the Research Data Assistance Center (ResDAC) for examples of how to appropriately suppress low counts\ [@cms_supression_policy].
:::

::: {.callout-important}
## Reporting Citations of Your Findings

Researchers are obligated to provide a list of citations for any publications or references to their research conducted using Cosmos. This list must be submitted to the Governing Council on a semi-annual basis.
:::

:::


## Valuable Links

::: {style="margin-left: 2rem"}
- [YNHH Learning Management System (LMS)](https://ynhh.certpointsystems.com/): Center Point Systems is where you can sign up for YNHH-created in-person or online training courses\ [@ynhhs_lms].

- [MyAPPs](https://myapps.ynhh.org): Find the portal to access the VDI through Citrix Workspace. Ensure your YNHH Duo is active for authentication and use your YU/YNHHS VPN if accessing from outside the Yale University or YNHHS network.

- [JDAT RBA SharePoint](https://yaleedu.sharepoint.com/:u:/r/sites/ResearcherBasicAccess/SitePages/CollabHome.aspx?csf=1&web=1&e=vQamcY): Find details about what RBA provides, along with supporting documents such as how-to guides for accessing the VDI, enrolling in Duo, links to training portals, and user guides for researchers\ [@jdat_sharepoint].

- [YNHH IT SharePoint](https://itstraining.ynhh.org): Find IT support materials, guides for navigating the Epic VDI software, notices of upcoming training events and system updates, and training requirements for researchers using Epic data. To access the Epic-specific documentation, click the blue "Training & Support Material" drop-down menu and select "Epic Training" to access the "Epic Training Portal"\ [@ynhhs_its_sharepoint].

- [JDAT Analytics Portal](https://jdat.ynhh.org/AnalyticsPortal): Submit a Helix request to have JDAT compile your research data. Additionally, access JDAT-created training for tools like Epic's SlicerDicer\ [@jdat_helix].

- [Epic UserWeb](https://userweb.epic.com/): Find Epic training materials, documentation, community discussions, support resources, updates, customization options, and access to your Epic profile\ [@epic_user_web].

- [Epic University](https://training.epic.com/): A subdomain of the UserWeb, find training opportunities and access to coursework necessary to obtain certificates, such as the Super User Badge, with courses organized by the Epic team. Learn more about advanced training options on our [Epic Cosmos](Epic-Cosmos.qmd) information webpage\ [@epic_training].

- [Galaxy](https://galaxy.epic.com/?#Home): A subdomain of the UserWeb, find comprehensive resource that provides documentation, announcements, best practices, and implementation guides related to the Epic dataset\ [@epic_galaxy].

:::


## Publications

```{r}
#| echo: false

# Fetch and format the citations.
citations <- get_citations(
  doi = c("10.1093/jamia/ocae098"),
  affiliation = c("Yale School of Public Health", "Yale School of Medicine", "Yale University"),
  highlight_names = c("Bhramar Mukherjee")
)

# Condition the output based on presence of articles.
citations_html <- make_pub_list(citations)

if(!any(str_detect(citations$citations, "No publications"))) {
  citation_length <- length(citations$citations)
  citation_date   <- str_match(citations$query, "\\((\\d{4}):(\\d{4})\\[dp\\]\\)")[,2]
} else {
  citation_length <- 0
  citation_date   <- as.numeric(format(Sys.Date(), "%Y")) - as.numeric(str_extract("No publications found in the last 25 years.", "[0-9]{2}"))
}

```


This section presents a selection of PubMed articles that utilize the dataset and are authored by individuals affiliated with the Yale School of Public Health. These articles are provided to inspire researchers and students to use the data in their own work.

::: {style="margin-left: 1rem"}

```{r}
#| echo: false
#| results: 'asis'

# Display formatted citations.
cat(citations_html)
```

```{=html}
<!-- Pagination buttons and scripting -->
<div class="citations-navigation">
  <button onclick="prevPage()">&#9664; Previous</button>
  <button onclick="nextPage()">Next &#9654;</button>
</div>

<script src="../../../Code/pagination.js"></script>
```

:::


## References
