---
title-block-banner: true
title: |
  Claude D. Pepper Older Americans</br>Independence Center (OAIC)
description: |
  </br>
  The Claude D. Pepper Older Americans Independence Centers (OAIC) Program is a network of 15 NIA-funded research centers established in 1991 to promote health and independence in older Americans. The OAIC Coordinating Center manages study registration, data access, and connects researchers with funding opportunities. Each center has its own research focus; Yale's Center focuses on multifactorial geriatric conditions and multimorbidity\ [@kitzman2024; @oaic_home].
include-in-header: ../../../_includes/feedback_popup.html
bibliography: BibTeX/oaic.bib
csl: ../../../Code/nature.csl
---

```{r}
#| echo: false

# Load in the functions.
source('../../../Code/PubMed Search_Functions.R')
```

<small>Updated: March 3<sup>rd</sup>, 2026</small>

## Overview

The [Claude D. Pepper Older Americans Independence Centers (OAIC) Program](https://www.peppercenter.org/) was established in 1991 through S.243 - Older Americans Act Reauthorization Amendments, signed by the 102<sup>nd</sup> Congress, which outlined how to develop, coordinate, and manage statewide programs that secure and maintain benefits and rights for older Americans\ [@oaic_home; @congress_s243]. The program honors US Representative Claude D. Pepper, who served in Congress for five decades championing healthcare reform and economic security for older Americans\ [@about_pepper]. His achievements included guaranteeing decent wages, expanding the National Institutes of Health (NIH), and establishing protections against age discrimination\ [@fsu_pepper].

Pepper's influence extended to sponsoring NIH institutes like the National Institute on Aging (NIA), which founded the OAIC program as one of its first Center-based initiatives\ [@kitzman2024; @fsu_pepper]. It has since expanded to 15 participating Centers, each specializing in specific research and administrative themes. Each Center has at least three cores: leadership and administrative, pilot/exploratory studies, and at least one research resource. Together, they support ongoing age-related research, produce specialized tools, and create training infrastructure for the next generation of investigators and leaders. For a complete breakdown of all centers and their research foci, see Table 1 of [Kitzman et al., 2025](https://pmc.ncbi.nlm.nih.gov/articles/PMC11090727/)\ [@kitzman2024].

The OAIC primary website links to Center-based resources, while its Coordinating Center (OAIC CC) manages study registration, stores limited numerical and structured text data from affiliated studies, and handles data access requests. Biospecimens are not managed by the OAIC CC; instead, study registrations provide primary contact information for those responsible for biospecimen storage and dissemination. It also connects researchers and early career investigators with seed funding opportunities and career development guidance in aging research.

[Yale's OAIC Center](https://medicine.yale.edu/internal-medicine/geriatrics/peppercenter/cores/biostatistics/) includes five cores: leadership and administrative, biostatistics, operations for data management, pilot and exploratory studies, and research education\ [@yale_oaic]. Its research domain concentrates on multifactorial geriatric conditions and multimorbidity in older adults. The following two resources were created with significant contributions from Yale's Center:

::: {style="margin-left: 2rem"}
- [**Geriatric Research Algorithms & Statistical Programs (GRASP):**](https://www.peppercenter.org/public/grasp.cfm) Curated collection of statistical analysis programs for aging research biostatisticians, providing downloadable code, sample data, resource links, and associated documents, developed by the OAICs at Yale, Duke, and Wake Forest Universities with NIA support\ [@oaic_grasp].

- [**Geriatric Research Instrument Library (GRIL):**](https://www.peppercenter.org/public/gril.cfm) Web-based repository for identifying, comparing, and selecting measurement instruments for gerontological research through domain navigation or search, with instrument descriptions and manuscript links. Developed by the OAICs at Yale, Duke, and Wake Forest Universities with NIA support\ [@oaic_gril].
:::


## Gaining Access

Affiliated studies submit data to the Pepper Study Registry (PSR) in two forms: metadata only or metadata with raw study data. This provides access to numerical and structured text data, but not images or physical samples. While biospecimen metadata is accessible in the PSR, dissemination of biospecimens themselves is handled by the original submitter, who can be reached through their listed primary contact\ [@oaic_psr].

Researchers can review the [Registry](https://www.peppercenter.org/public/dspPSR.cfm#!) and request access through the OAIC's [member page](https://www.peppercenter.org/dspLogin.cfm)\ [@oaic_registry].

## Relevant Links

::: {style="margin-left: 2rem"}
- [Pepper Study Registry (PSR)](https://www.peppercenter.org/public/dspPSR.cfm): Find information about how to submit qualifying research to the PSR, browse currently available studies and primary contacts, and request access to data\ [@oaic_registry].

- [Yale Claude D. Pepper Older Americans Independence Center](https://medicine.yale.edu/internal-medicine/geriatrics/peppercenter/): Find the landing page for Yale's OAIC\ [@yale_oaic].

- [Geriatric Research Algorithms & Statistical Programs (GRASP)](https://www.peppercenter.org/public/grasp.cfm): Find statistical analysis programs for aging research biostatisticians with downloadable code, sample data, and resource links. Developed by the OAICs at Yale, Duke, and Wake Forest Universities\ [@oaic_grasp].

- [Geriatric Research Instrument Library (GRIL)](https://www.peppercenter.org/public/gril.cfm): Find a repository for identifying, comparing, and selecting measurement instruments for gerontological research with instrument descriptions and manuscript links. Developed by the OAICs at Yale, Duke, and Wake Forest Universities\ [@oaic_gril].

- [Researcher Career Roadmap](https://www.peppercenter.org/public/career_roadmap.cfm): Find information about career support for age-related researchers at various stages, organized by education level and including federal and foundation opportunities\ [@oaic_roadmap].

- [Events of Interest](https://www.peppercenter.org/public/announcements.cfm?select_category=Funding%20Opportunity): Find upcoming events, workshops, conferences, symposiums, jobs, and funding deadlines\ [@oaic_events].

- [Video Showcase](https://vimeo.com/showcase/10347393): Find recorded annual meetings and information sessions from the OAIC CC\ [@oaic_videos].

- [Kitzman et al., 2025. The Pepper Older Americans Independence Centers: an NIA-sponsored Program for Improving Physical Function among Older Persons](https://pmc.ncbi.nlm.nih.gov/articles/PMC11090727/): Find the article discussing OAIC's structure, its role in age-related research, and participating centers with their research focus areas\ [@kitzman2024].
:::


## Publications

```{r}
#| echo: false

# Fetch and format the citations.
citations <- get_citations(
  doi = c("10.1177/26335565221081200", "10.1172/jci.insight.168102"),
  affiliation = c("Yale School of Public Health", "Yale School of Medicine", "Yale University"),
  highlight_names = c("Bhramar Mukherjee")
)

# Condition the output based on presence of articles.
citations_html <- make_pub_list(citations)

if(!any(str_detect(citations$citations, "No publications"))) {
  citation_length <- length(citations$citations)
  citation_date   <- str_match(citations$query, "\\((\\d{4}):(\\d{4})\\[dp\\]\\)")[,2]
} else {
  citation_length <- 0
  citation_date   <- as.numeric(format(Sys.Date(), "%Y")) - as.numeric(str_extract("No publications found in the last 25 years.", "[0-9]{2}"))
}

```


This section presents a selection of PubMed articles that utilize the dataset and are authored by individuals affiliated with the Yale University. These articles are provided to inspire researchers and students to use the data in their own work.

::: {style="margin-left: 1rem"}

```{r}
#| echo: false
#| results: 'asis'

# Display formatted citations.
cat(citations_html)
```

```{=html}
<!-- Pagination buttons and scripting -->
<div class="citations-navigation">
  <button onclick="prevPage()">&#9664; Previous</button>
  <button onclick="nextPage()">Next &#9654;</button>
</div>

<script src="../../../Code/pagination.js"></script>
```

:::

## References
