---
title-block-banner: true
title: |
  data.gov
description: |
  </br>
  Data.gov is the U.S. federal government's official open data website, launched in 2009 to enhance government transparency and accountability by providing free public access to federal datasets. Operating under the OPEN Government Data Act of 2018, it mandates federal agencies to publish their data online in standardized, machine-readable formats. The platform, maintained by the General Services Administration (GSA), also collaborates with state, local, and international sources to create a comprehensive data catalog\ [@user_guide]. Its open-source foundation enables adaptation by governments worldwide, with the source code available on the GSA's GitHub repository\ [@github].
include-in-header: ../../../_includes/feedback_popup.html
bibliography: BibTeX/data_gov_references.bib
csl: ../../../Code/nature.csl
---

```{r}
#| echo: false

# Load in the functions.
source('../../../Code/PubMed Search_Functions.R')
```

<small>Updated: January 20<sup>th</sup>, 2026</small>

## Overview

Data.gov is the U.S. federal government's official open data website, designed to make government more open and accountable by providing free public access to federal datasets. Originally launched in 2009 following a presidential directive for greater transparency, the platform operates under the OPEN Government Data Act of 2018, which legally requires federal agencies to publish their data online in standardized, machine-readable formats. The site promotes citizen participation in government, economic development, and informed decision-making across both public and private sectors, with the General Services Administration (GSA) responsible for maintaining the central catalog\ [@open_gov; @user_guide].

Data.gov extends beyond federal data by collaborating with state, local, and international sources to build a comprehensive multi-level government catalog. The platform's open source foundation allows governments worldwide to adapt its code for their own initiatives. The source code is available on the GSA [GitHub](https://github.com/GSA/data.gov) repository, and a complete listing of international and U.S. open data sites can be found on the [Open Data in the United States and Internationally](https://data.gov/open-gov/#open-data-in-the-united-states-and-internationally) page\ [@github; @github]. Federal agencies currently represented include\ [@catalog]:

::: {style="margin-left: 2rem"}
- **Health & Medical Agencies:** National Institutes of Health (NIH), Centers for Disease Control and Prevention (CDC), Centers for Medicare and Medicaid Services (CMS), Substance Abuse and Mental Health Services Administration (SAMHSA), Veterans Health Administration (VHA), Administration for Children and Families (ACF), Department of Health and Human Services (HHS)

- **Environmental & Natural Resources:** United States Geological Survey (USGS), Environmental Protection Agency (EPA), National Park Service (NPS), Bureau of Reclamation, Bureau of Land Management (BLM), United States Fish and Wildlife Service (USFWS), Forest Service, Bureau of Safety and Environmental Enforcement (BSEE)

- **Space & Scientific Research:** National Aeronautics and Space Administration (NASA), National Institute of Standards and Technology (NIST), Nuclear Regulatory Commission (NRC), Office of Science and Technology Policy (OSTP)

- **Demographics & Social Statistics:** Bureau of the Census, Social Security Administration (SSA)

- **Agriculture & Food:** Agricultural Research Service (ARS), Economic Research Service (ERS), Food and Nutrition Service (FNS), Farm Service Agency (FSA)

- **Transportation & Infrastructure:** Department of Transportation (DOT), Federal Highway Administration (FHWA)

- **Justice & Law Enforcement:** Office of Justice Programs (OJP)

- **Education:** Institute of Education Sciences (IES), Department of Education, Office of English Language Acquisition, Office of Postsecondary Education

- **Energy:** Energy Programs

- **Financial & Economic:** Federal Financing Bank, Financing Vehicles and the Board of Governors of the Federal Reserve, Federal Trade Commission (FTC)

- **Veterans Affairs:** Department of Veterans Affairs (VA), Departmental Administration

- **International & Trade:** Department of State, International Trade Commission (ITC)

- **Housing & Urban Development:** Department of Housing and Urban Development (HUD)

- **Employment & Personnel:** Office of Personnel Management (OPM), Equal Employment Opportunity Commission (EEOC)
:::

## Gaining Access

All data are freely accessible for direct download or online visualization through the platform. Simply visit the [Data Catalog](https://catalog.data.gov/dataset) page to download directly\ [@catalog].

::: {.callout-important}
Terms of service and licenses for non-federal datasets accessible on data.gov may vary from those of federal agency sources.
:::

## Relevant Links

::: {style="margin-left: 2rem"}
- [Data Catalog](https://catalog.data.gov/dataset): Search through the complete data catalog by topic, categories, etc. See how many datasets are applied to each listing filter type\ [@catalog].

- [User Guide](https://data.gov/user-guide/): Find information on how to effectively search the catalog of data, discover the contributors to this database, and understand what you might find on each data page, including any applicable terms of service and licenses\ [@user_guide].

- [Open Government](https://data.gov/open-gov/): Find information about federal open data site requirements and discover other international and U.S. open data sites\ [@open_gov].

- [Resources.data.gov](resources.data.gov): Find policies, tools, case studies, and other resources to support federal data governance, management, exchange, and use and help agencies adopt open data practices across government\ [@resources].

- [GSA/data.gov](https://github.com/GSA/data.gov): Find Data.gov platform code, GitHub Actions templates, and development resources at the main Data.gov Platform repository, which houses reusable code and documentation for the static site (www.data.gov), catalog (catalog.data.gov), and inventory (inventory.data.gov) applications\ [@github].
:::


## Publications

```{r}
#| echo: false

# Fetch and format the citations.
citations <- get_citations(
  doi = c("10.1001/jamainternmed.2024.5932", "10.1097/XCS.0000000000000429",
          "10.1093/cid/ciae427"),
  affiliation = c("Yale School of Public Health", "Yale School of Medicine", "Yale University"),
  highlight_names = c("Bhramar Mukherjee")
)

# Condition the output based on presence of articles.
citations_html <- make_pub_list(citations)

if(!any(str_detect(citations$citations, "No publications"))) {
  citation_length <- length(citations$citations)
  citation_date   <- str_match(citations$query, "\\((\\d{4}):(\\d{4})\\[dp\\]\\)")[,2]
} else {
  citation_length <- 0
  citation_date   <- as.numeric(format(Sys.Date(), "%Y")) - as.numeric(str_extract("No publications found in the last 25 years.", "[0-9]{2}"))
}

```


This section presents a selection of PubMed articles that utilize the dataset and are authored by individuals affiliated with the Yale University. These articles are provided to inspire researchers and students to use the data in their own work.

::: {style="margin-left: 1rem"}

```{r}
#| echo: false
#| results: 'asis'

# Display formatted citations.
cat(citations_html)
```

```{=html}
<!-- Pagination buttons and scripting -->
<div class="citations-navigation">
  <button onclick="prevPage()">&#9664; Previous</button>
  <button onclick="nextPage()">Next &#9654;</button>
</div>

<script src="../../../Code/pagination.js"></script>
```

:::

## References
