---
title-block-banner: true
title: |
  The Health Attribution Library
description: |
  </br>
  The Health Attribution Library is a curated collection of scientific publications that conduct rigorous end-to-end attribution studies linking human-caused climate change to specific health impacts. Created in 2023 through Wellcome Trust-funded research that found only 13 qualifying studies among nearly 4,000 reviewed papers, the library serves as an ongoing resource to support scientific research, journalism, litigation, and policy work requiring evidence-based attribution of health impacts to climate change\ [@attributes_main; @carlson2024].
include-in-header: ../../../_includes/feedback_popup.html
bibliography: BibTeX/yale_freely_references.bib
csl: ../../../Code/nature.csl
---

```{r}
#| echo: false

# Load in the functions.
source('../../../Code/PubMed Search_Functions.R')
```

<small>Updated: January 20<sup>th</sup>, 2026</small>

## Overview

The Health Attribution Library is a curation of scientific publications conducting end-to-end attribution studies on the health impacts of anthropogenic climate change. End-to-end studies, also called "one-step" studies, use climate models to estimate health impacts with and without human-caused climate change. While these are labor- and computationally intensive study designs, they yield more information about uncertainty along the steps of the analysis\ [@attributes_main; @carlson2024].

The library was created as an initial literature review in 2023, contracted by The Wellcome Trust to [Colin Carlson, PhD](https://ysph.yale.edu/profile/colin-carlson/) (Yale University, USA) and Christopher Trisos, PhD (University of Cape Town, South Africa) as part of an expert panel exploring issues in the field of health impact attribution. Their findings were published in _Detection & Attribution of Climate Change Impacts on Human Health_ with The Wellcome Trust in 2024\ [@carlson2024].

Their research revealed a significant gap in the field: of nearly 4,000 peer-reviewed studies reviewed, only 13 studies since 2013 rigorously tackled health impact attribution. Most studies focused on direct effects of climate events or failed to evaluate climate-sensitive disease burdens, such as zoonotic, tick-borne, and water-borne diseases, to elucidate climate-attributable health impacts\ [@carlson2024].

The library is intended as an ongoing curation of qualifying health attribution research to support scientific research, journalism, public interest work, litigation, and policy negotiations seeking to rigorously answer whether specific health impacts can be attributed to human-caused climate change\ [@carlson2024].


## Gaining Access

All papers are freely accessible for direct download or online visualization through the platform. Access them by visiting the [Database](https://www.healthattribution.org/database) page\ [@attributes_database].

## Relevant Links

::: {style="margin-left: 2rem"}
- [Database](https://www.healthattribution.org/database): Access and download papers directly\ [@attributes_database].

- [Carlson et al. 2024. Detection & Attribution of Climate Change Impacts on Human Health](https://pmc.ncbi.nlm.nih.gov/articles/PMC10409981/): Find the article introducing health impact attribution research, including its conceptual framework, scientific goals, current progress, and future research needs\ [@carlson2024].
:::


## Publications

```{r}
#| echo: false

# # Fetch and format the citations.
# citations <- get_citations(
#   doi = c("10.1016/j.urology.2022.07.012"),
#   affiliation = c("Yale School of Public Health", "Yale School of Medicine", "Yale University"),
#   highlight_names = c("Bhramar Mukherjee")
# )
# 
# # Condition the output based on presence of articles.
# citations_html <- make_pub_list(citations)
# 
# if(!any(str_detect(citations$citations, "No publications"))) {
#   citation_length <- length(citations$citations)
#   citation_date   <- str_match(citations$query, "\\((\\d{4}):(\\d{4})\\[dp\\]\\)")[,2]
# } else {
#   citation_length <- 0
#   citation_date   <- as.numeric(format(Sys.Date(), "%Y")) - as.numeric(str_extract("No publications found in the last 25 years.", "[0-9]{2}"))
# }

```


This section presents a selection of PubMed articles that utilize the dataset and are authored by individuals affiliated with the Yale University. These articles are provided to inspire researchers and students to use the data in their own work.

::: {style="margin-left: 1rem"}

```{r}
#| echo: false
#| results: 'asis'

# # Display formatted citations.
# cat(citations_html)
```

```{=html}
<!-- Pagination buttons and scripting -->
<div class="citations-navigation">
  <button onclick="prevPage()">&#9664; Previous</button>
  <button onclick="nextPage()">Next &#9654;</button>
</div>

<script src="../../../Code/pagination.js"></script>
```

:::

## References
