---
title-block-banner: true
title: |
  National Hospital Care Surveys (NHCS)
description: |
  </br>
  The National Hospital Care Surveys (NHCS) collect and disseminate data to provide comprehensive understanding of hospital healthcare services in the United States. The data includes patient demographics, diagnoses, treatments, procedures, and outcomes from emergency departments, inpatient units, office-based providers, health centers, and residential care communities, and is accessible through reports and data files to inform health policy decisions, monitor care quality, and examine healthcare disparities\ [@nhcs_all_about].
include-in-header: ../../../_includes/feedback_popup.html
bibliography: BibTeX/all_nchs_references.bib
csl: ../../../Code/nature.csl
---

```{r}
#| echo: false

# Load in the functions.
source('../../../Code/PubMed Search_Functions.R')
```

<small>Updated: December 22<sup>th</sup>, 2025</small>

## Overview

The National Hospital Care Surveys (NHCS) are comprehensive national programs conducted by the National Center for Health Statistics (NCHS), which is part of the Centers for Disease Control and Prevention (CDC). These surveys collect critical data on U.S. hospital healthcare services to inform health policy decisions, public health initiatives, and healthcare service improvements by providing information about organizations and providers delivering health care, the services they provide, and the patients they serve\ [@nhcs_all_about].

This provider-based data comes directly from healthcare professionals and facility operators and offers nationally representative insights into U.S. health providers, services, patients, and healthcare facilities. The NHCS surveys capture healthcare information from six key areas\ [@nhcs_all_about; @nhcs_all_current]:

::: {style="margin-left: 2rem"}
- **Hospital emergency departments:** Patient demographics, presenting complaints, triage classifications, diagnoses, treatments provided, procedures performed, medications administered, length of stay, disposition decisions, and clinical outcomes including transfers and readmissions.

- **Hospital inpatient departments:** Patient demographics, admission sources and types, primary and secondary diagnoses, surgical and non-surgical procedures, length of stay, complications, discharge destinations, readmission rates, and resource utilization including ICU stays.

- **Office-based providers:** Outpatient visits including patient characteristics, treatments, medications prescribed, diagnostic tests, and follow-up care plans.

- **Health centers:** Community-based primary care services, patient demographics, visit types, preventive care, and chronic disease management.

- **Adult day centers and residential care communities:** Care services, participant demographics, functional status, facility characteristics, and staffing information.
:::

Through its survey programs, the NHCS examines healthcare resource utilization, quality and safety measures, costs, disease-specific care patterns, service disparities across populations, and adoption of treatments and technologies like electronic health records (EHR). To ensure data accuracy, multiple collection methodologies are employed, including medical record abstraction by trained personnel, electronic data extraction from EHRs, and surveys distributed to healthcare providers and patients. Each survey maintains stable core topics to enable reliable trend monitoring across healthcare settings over time\ [@nhcs_all_about].

## Gaining Access

NHCS consists of five surveys, each detailed on individual information pages\ [@nhcs_all_current]. Access information for each program's data is available on their respective pages from our [Dataset](../../datasets.qmd) listing page.

::: {style="margin-left: 2rem"}
- **Ambulatory Medical Care Survey (NAMCS):** Collects information on office visits for patients receiving same-day care, not admitted to hospitals, including data on telemedicine usage, electronic health record implementation, and healthcare delivery patterns.

- **Electronic Health Records Survey (NEHRS):** Provides data on healthcare offices' adoption of EHR systems, electronic health information exchange, and EHR-related administrative burden. This program tracks participation in the Centers for Medicare and Medicaid Services' Promoting Interoperability Programs.

- **Hospital Ambulatory Medical Care Survey (NHAMCS):** Combines data on patients receiving hospital-based ambulatory (same-day) care services in emergency and outpatient departments without hospital admission. This survey was conducted annually through 2022 and transitioned from manual medical record collection to electronic health record extraction as EHRs became widely adopted.

- **Hospital Care Survey (NHCS):** Collects hospital inpatient discharge and emergency department visit data. Hospitals also provide summary information about patient visits for the calendar year through supplemental interviews.

- **Post-acute and Long-term Care Study (NPALS):** Collects and analyzes data from providers delivering post-acute care (following urgent medical treatment) and long-term care (for patients unable to care for themselves over extended periods). It utilizes existing Centers for Medicare and Medicaid Services data for five sectors and directly collects data from adult day services centers and residential care communities.
:::

Prior to these current surveys, earlier iterations existed that either evolved into the surveys listed above or were discontinued. Data from these earlier surveys remain accessible and can sometimes be combined with relevant modern iterations. Below are the inactive NHCS surveys, with notes indicating which evolved into current surveys\ [@nhcs_all_past].

Unlike the currently maintained surveys, these discontinued surveys do not have dedicated more info pages on this site. Researchers are encouraged to explore details about these surveys on their archived CDC pages.

::: {style="margin-left: 2rem"}
- [**Home and Hospice Care Survey (NHHCS):**](https://archive.cdc.gov/www_cdc_gov/nchs/nhhcs/index.htm) Gathered comprehensive data on home health and hospice care agencies, focusing on three key areas: patient characteristics, services provided, and agency details. The survey ran from 1992 to 2007, with data collection occurring in 1992, 1993, 1994, 1996, 1998, 2000, and 2007. In 2012, it transitioned into NPALS, which was previously known as NSLTCP\ [@npals_about_alt; @nhhcs_about_archived].

- [**Hospital Discharge Survey (NHDS):**](https://archive.cdc.gov/www_cdc_gov/nchs/nhds/index.htm) Compiled nationally representative data on inpatients discharges from nonfederal, short-stay hospitals in the United States from 1965 to 2010. The survey focused on general and children's hospitals with average stays under 30 days, excluding veterans' hospitals, institutions serving confined populations, and facilities without inpatient capabilities. At its largest, the survey included 422 hospitals in 2007. The NHDS's inpatient data collection role has since been assumed by the NHCS survey, while NHAMCS now handles emergency department, outpatient department, and ambulatory surgery center data\ [@nhds_about_archived].

- [**Nursing Home Survey (NNHS):**](https://archive.cdc.gov/www_cdc_gov/nchs/nnhs/index.htm) Nationally representative surveys of nursing homes, examining facilities, services, staff, and residents. Operating from 1973 to 2004 (with surveys conducted in 1973-74, 1977, 1985, 1995, 1997, 1999, and 2004), each survey collected core information about nursing homes while focusing on different topics. Participating facilities had at least three beds and were either certified by Medicare/Medicaid or state-licensed. In 2012, the NNHS evolved into NPALS, previously known as NSLTCP. NPALS broadened its scope to cover adult day services centers, residential care communities, nursing homes, home health agencies, and hospice agencies, using CMS administrative data for nursing home information\ [@npals_about_alt; @nnhs_about_archived].

- [**Survey of Ambulatory Surgery (NSAS):**](https://archive.cdc.gov/www_cdc_gov/nchs/nsas/index.htm) Collected data on ambulatory surgery (same-day outpatient surgeries) visits from hospitals and freestanding surgery centers during 1994-1996 and 2006, serving as the sole source of nationally representative ambulatory surgery data. In 2009, NHAMCS took over ambulatory surgery data collection, incorporating hospital data into its ongoing emergency department and outpatient department surveys (since 1992). NHAMCS briefly expanded to cover ambulatory surgery centers from 2010-2012 before discontinuing this in 2013 and maintaining only hospital-based collection through 2017\ [@nsas_about_archived].

- [**Survey of Residential Care Facilities (NSRCF):**](https://archive.cdc.gov/www_cdc_gov/nchs/nsrcf/index.htm) Conducted in 2010, this was the first-ever comprehensive survey of residential care providers. It covered residential care facilities, assisted living residences, board and care homes, congregate care, enriched housing programs, homes for the aged, personal care homes, and shared housing establishments that are licensed, registered, listed, certified, or otherwise regulated by a state. In 2012, NSRCF evolved into NPALS (previously known as NSLTCP), which covers adult day services centers, residential care communities, nursing homes, home health agencies, and hospice agencies. NPALS continues collecting residential care community data alongside these other long-term care providers\ [@npals_about_alt; @nsrcf_about_archived].
:::

## Valuable Links

::: {style="margin-left: 2rem"}
- [Current Surveys](https://www.cdc.gov/nchs/healthcare-surveys/about/surveys.html): Find a comprehensive listing of all active surveys conducted under the NHCS program\ [@nhcs_all_current].

- [Past Surveys](https://www.cdc.gov/nchs/healthcare-surveys/about/past-surveys.html): Find a comprehensive listing of all surveys previously conducted under the NHCS program\ [@nhcs_all_past].
:::


## Publications

```{r}
#| echo: false

# Fetch and format the citations.
citations <- get_citations(
  doi = c("10.5055/jom.0862", "10.1007/s11606-021-06668-x", "10.1007/s11606-022-07907-5",
          "10.4088/PCC.24br03776"),
  affiliation = c("Yale School of Public Health", "Yale School of Medicine", "Yale University"),
  highlight_names = c("Bhramar Mukherjee")
)

# Condition the output based on presence of articles.
citations_html <- make_pub_list(citations)

if(!any(str_detect(citations$citations, "No publications"))) {
  citation_length <- length(citations$citations)
  citation_date   <- str_match(citations$query, "\\((\\d{4}):(\\d{4})\\[dp\\]\\)")[,2]
} else {
  citation_length <- 0
  citation_date   <- as.numeric(format(Sys.Date(), "%Y")) - as.numeric(str_extract("No publications found in the last 25 years.", "[0-9]{2}"))
}

```


This section presents a selection of PubMed articles that utilize the dataset and are authored by individuals affiliated with the Yale School of Public Health. These articles are provided to inspire researchers and students to use the data in their own work.

::: {style="margin-left: 1rem"}

```{r}
#| echo: false
#| results: 'asis'

# Display formatted citations.
cat(citations_html)
```

```{=html}
<!-- Pagination buttons and scripting -->
<div class="citations-navigation">
  <button onclick="prevPage()">&#9664; Previous</button>
  <button onclick="nextPage()">Next &#9654;</button>
</div>

<script src="../../../Code/pagination.js"></script>
```

:::

## References
