---
title-block-banner: true
title: |
  nextstrain
description: |
  </br>
  The Surveillance, Epidemiology, and End Results Program (SEER) has provided vital cancer statistics since 1973, now covering 45.9% of the US population and recording around 17 million cases. SEER data includes demographics, tumor characteristics, treatments, and outcomes from various registries, providing national representation of different racial and ethnic groups\ [@about_seer; @program_overview].
  </br>
  </br>
  Researchers can access SEER Research Data, SEER Research Plus and NCCR Data, Supplementary Datasets, and Specialized Datasets, each offering varying levels of detail. SEER also provides software tools like SEER\*Explorer, NCCR\*Explorer, SEER\*Stat, SEERPrep, and SEER\*DMS for efficient data access and analysis\ [@data_apps; @explorer_tools; @methods_tools; @data_software].
include-in-header: ../../../_includes/feedback_popup.html
bibliography: BibTeX/yale_freely_references.bib
csl: ../../../Code/nature.csl
---

```{r}
#| echo: false

# Load in the functions.
source('../../../Code/PubMed Search_Functions.R')
```

<small>Updated: October 21<sup>th</sup>, 2025</small>

## Overview



The types of data products available include\ [@data_apps; @data_variables; @data_comparisons; @us_data]:

::: {style="margin-left: 2rem"}
- **SEER Research Data:** The standard SEER data, which provides information on cancer incidence and survival from population-based cancer registries. The data encompasses patient demographics, primary tumor site, tumor morphology, stage at diagnosis, extent and history of the disease, initial treatments, and survival outcomes.
:::


## Gaining Access

::: {.panel-tabset}

### Research Data

#### Do I Qualify?

#### Typical Timeline

#### Step-by-Step Guide

:::

## Relevant Links

::: {style="margin-left: 2rem"}
- [Documentation for SEER Data](https://seer.cancer.gov/data-software/documentation/): Find details about SEER databases, the NCCR databases, and variable definitions available in SEER*Stat, including record numbers, U.S. population coverage, geographic regions, and SEER variable dictionary\ [@data_docs].
:::


## Publications

```{r}
#| echo: false

# Fetch and format the citations.
citations <- get_citations(
  doi = c("10.1016/j.urology.2022.07.012"),
  affiliation = c("Yale School of Public Health", "Yale School of Medicine", "Yale University"),
  highlight_names = c("Bhramar Mukherjee")
)

# Condition the output based on presence of articles.
citations_html <- make_pub_list(citations)

if(!any(str_detect(citations$citations, "No publications"))) {
  citation_length <- length(citations$citations)
  citation_date   <- str_match(citations$query, "\\((\\d{4}):(\\d{4})\\[dp\\]\\)")[,2]
} else {
  citation_length <- 0
  citation_date   <- as.numeric(format(Sys.Date(), "%Y")) - as.numeric(str_extract("No publications found in the last 25 years.", "[0-9]{2}"))
}

```


This section presents a selection of PubMed articles that utilize the dataset and are authored by individuals affiliated with the Yale University. These articles are provided to inspire researchers and students to use the data in their own work.

::: {style="margin-left: 1rem"}

```{r}
#| echo: false
#| results: 'asis'

# Display formatted citations.
cat(citations_html)
```

```{=html}
<!-- Pagination buttons and scripting -->
<div class="citations-navigation">
  <button onclick="prevPage()">&#9664; Previous</button>
  <button onclick="nextPage()">Next &#9654;</button>
</div>

<script src="../../../Code/pagination.js"></script>
```

:::

## References
