---
title-block-banner: true
title: |
  Surveillance, Epidemiology, and</br>End Results Program (SEER)
description: |
  </br>
  The Surveillance, Epidemiology, and End Results Program (SEER) has provided vital cancer statistics since 1973, now covering 45.9% of the US population and recording around 17 million cases. SEER data includes demographics, tumor characteristics, treatments, and outcomes from various registries, providing national representation of different racial and ethnic groups\ [@about_seer; @program_overview].
  </br>
  </br>
  Researchers can access SEER Research Data, SEER Research Plus and NCCR Data, Supplementary Datasets, and Specialized Datasets, each offering varying levels of detail. SEER also provides software tools like SEER\*Explorer, NCCR\*Explorer, SEER\*Stat, SEERPrep, and SEER\*DMS for efficient data access and analysis\ [@data_apps; @explorer_tools; @methods_tools; @data_software].
include-in-header: ../../../_includes/feedback_popup.html
bibliography: BibTeX/seer_references.bib
csl: ../../../Code/nature.csl
---

```{r}
#| echo: false

# Load in the functions.
source('../../../Code/PubMed Search_Functions.R')
```

<small>Updated: October 21<sup>th</sup>, 2025</small>

## Overview

The Surveillance, Epidemiology, and End Results Program (SEER) is supported by the National Cancer Institute’s (NCI) Surveillance Research Program (SRP), which provides national leadership in cancer surveillance science and offers the tools and expertise necessary for collecting, analyzing, interpreting, and disseminating reliable population-based statistics. Since 1973, SEER has provided vital cancer statistics, growing to cover 45.9% of the US population across 22 geographic areas with its data on cancer incidence and survival. To date, they have recorded as many as approximately 17 million cases in some of their registries, almost 16 million of which represent malignant cases\ [@about_seer; @program_overview].

SEER data includes patient demographics, tumor characteristics, initial treatments, and survival outcomes, offering significant representation of various racial and ethnic groups across the nation\ [@data_variables; @program_overview]. The data is collected from participating cancer registries, which can be either population-based or hospital-based, with special registries focusing on specific cancers\ [@registries].

The types of data products available include\ [@data_apps; @data_variables; @data_comparisons; @us_data]:

::: {style="margin-left: 2rem"}
- **SEER Research Data:** The standard SEER data, which provides information on cancer incidence and survival from population-based cancer registries. The data encompasses patient demographics, primary tumor site, tumor morphology, stage at diagnosis, extent and history of the disease, initial treatments, and survival outcomes.

- **SEER Research Plus and NCCR Data:** This expanded category includes all the SEER Research Data as well as additional variables and detailed information not included in the standard dataset, such as geography, month in addition to year of diagnosis, and other demographic fields. It also encompasses data from the National Childhood Cancer Registry (NCCR), offering a more comprehensive dataset for in-depth research, including more variables related to specific cancer types and treatments.

- **Supplementary Datasets:** Data commonly used with to SEER or NCCR, including U.S. Mortality, U.S. Populations, Standard Populations, County/Tract Attributes, Expected Survival, and SEER Linked Databases. The Linked Databases include SEER-Medicare, SEER-Medicare Health Outcomes Survey (SEER-MHOS), SEER-Medicaid, and SEER-Consumer Assessment of Healthcare Providers and Systems (SEER-CAHPS).

- **Specialized Datasets:** Specialized Datasets include additional variables not found in the standard SEER Research Data and require access to the latest SEER Research Plus Data. Available databases cover various specific cancers and attributes, such as Pharyngeal Cancer with HPV Status, Prostate Cancer with additional fields, and Incidence Data with Census Tract Attributes, with some requiring NCI Central IRB approval\ [@specialized_data_login].
:::

The types of data collected and represented in these include\ [@data_variables]: 

::: {style="margin-left: 2rem"}
- **Demographic Data:** Basic demographics like age, gender, race, socioeconomic status, family characteristics, and geographic location information such as state and urbanicity.

- **Survival Outcomes:** Information on the patient’s survival status and follow-up data, including the length of survival and cause of death.

- **Tumor Characteristics:**  Details about the primary tumor site, morphology, stage at diagnosis, extent and history of the disease, and other relevant attributes.

- **Initial Treatments:** Data on the first course of treatment administered to patients, including surgery, radiation, and chemotherapy.
:::

In addition to its data, SEER has created software tools and webpages designed to simplify data access, analysis, and dissemination. They also interface with SEER data and other cancer-related databases for facilitating data retrieval and analyses\ [@explorer_tools; @methods_tools; @data_software].

::: {style="margin-left: 2rem"}
- **SEER\*Explorer:** A webpage that provides SEER cancer summary statistics by race, age, sex, stage, and cancer subtypes, covering 46% of the U.S. population.

- **NCCR\*Explorer:** A webpage that provides childhood, adolescent and young adult cancer summary statistics from National Childhood Cancer Registry (NCCR) data for ages 0-39 using International Classification of Childhood Cancer (ICCC).

- **SEER\*Stat:** A computer software for analyzing SEER data and other cancer-related databases, specializing in endopoints about incidence rates, survival, and prevelance.

- **SEERPrep:** Converts user-supplied ASCII text data files to SEERStat database format.

- **SEER Data Management System (SEER\*DMS):** Supports all core cancer registry functions including importing data, editing, linkage, consolidation, and reporting.
:::


## Gaining Access

::: {.panel-tabset}

### Research Data

#### Do I Qualify?

All persons with a valid email address are allowed to access the data\ [@data_apps].

#### Typical Timeline

After completing the application steps below, the SEER program will process your request and provide you with a SEER\*Stat account, if approved. This account can be used to access data through SEER\*Stat.


#### Step-by-Step Guide

Complete details can be found on the [How to Request Access to SEER Data](https://seer.cancer.gov/data/access.html) webpage and are summarized below\ [@access_policy]:

::: {style="margin-left: 2rem"}
1. Register in the [Data Request System](https://seerdataaccess.cancer.gov/seer-data-access) with your email and complete the application form.

2. Acknowledge the SEER Research Data User Agreement and SEER Treatment Data Limitations documents and submit. Note that these documents are updated periodically.

3. Verify your email address after submission.
:::

#### Download and Install SEER\*Stat

SEER\*Stat is a PC software application that offers an intuitive user interface for analyzing SEER data and other cancer-related databases. An internet connection is required to use it\ [@install_seer_stat].

::: {style="margin-left: 2rem"}
1. Download the current version of Microsoft Windows from the [SEER\*Stat Installation](https://seer.cancer.gov/seerstat/software/) webpage.

2. Log into SEER*Stat using the email address you used to request the data.

3. Set your password by navigating to the "Client-Server Set/Change Password" link in the profile menu and entering your credentials.
:::


### Research Plus and NCCR Data

#### Do I Qualify?

Researchers need to have an eRA Commons or HHS account to access the data. Starting April 4, 2025, the NIH will prohibit institutions located in countries of concern—including China (including Hong Kong and Macau), Russia, Iran, North Korea, Cuba, and Venezuela—from accessing NIH Controlled-Access Data Repositories and associated data.

#### Typical Timeline

The timeline varies depending on verification with HHS or Login.gov, and whether you already have an existing Login.gov account. It can take about two days for the SEER Data Request System to register that Login.gov is linked with eRA. After submitting your application, the SEER program will process requests within two business days. If approved, you will receive an email with your SEER*Stat account details and information on how to access data and download the software\ [@access_policy].

#### Step-by-Step Guide

Complete details can be found on the [How to Request Access to SEER Data](https://seer.cancer.gov/data/access.html) webpage and are summarized below:

::: {style="margin-left: 2rem"}
1. Log into [Data Request System](https://seerdataaccess.cancer.gov/seer-data-access) using one of the following methods:

::: {style="margin-left: 1rem"}
**HHS:** Using your PIV card.

**eRA Commons:**

::: {style="margin-left: 2rem"}
a. If you do not have an eRA Commons account, you will need to create one in coordination with your institution.

b. Confirm that your eRA account is associated with the institutional email address you want to use to request data (e.g., .edu, .gov, .org, or work email address).

c. Link your eRA account with Login.gov by logging into the [eRA Commons](https://public.era.nih.gov/commons/) website using the "Login with Login.gov" option. If you do not have a Login.gov account, follow the instructions to create one. If logging in this way takes you directly to the eRA profile page, your accounts are already linked; otherwise, follow the on-screen prompts.

:::{.callout-note}
It may take up to two days for the SEER Data Request System to register that the accounts are linked. Refer to the [FAQ](https://seer.cancer.gov/data/faqs.html#accounts-linked) page for additional details if you encounter problems linking the accounts\ [@faq].
:::
:::
:::

2. Complete the application form in the Data Request System and provide the Signing Official's contact information. If you have trouble finding this information, refer to the [FAQ](https://seer.cancer.gov/data/faqs.html#signing-official) guide\ [@faq].

3. Acknowledge the SEER Research Data Use Agreement, SEER Treatment Data Limitations, and National Childhood Cancer Registry (NCCR) Data Use Agreement and submit. Note that these documents are updated periodically.
:::

#### Download and Install SEER\*Stat

SEER\*Stat is a PC software application that offers an intuitive user interface for analyzing SEER data and other cancer-related databases. An internet connection is required to use it\ [@install_seer_stat].

::: {style="margin-left: 2rem"}
1. Download the current version of Microsoft Windows from the [SEER\*Stat Installation](https://seer.cancer.gov/seerstat/software/) webpage.

2. Log into SEER*Stat using the email address you used to request the data.

3. Set your password by navigating to the "Client-Server Set/Change Password" link in the profile menu and entering your credentials.
:::


### Specialized Datasets

#### Do I Qualify?

You must already have access to the Research Plus and NCCR Data before you can request specialized data.

#### Typical Timeline

Application reviews typically take at most one week for additional information or clarification. Data Access Committee (DAC) members meet bi-weekly and require three attending members to vote for approval of the request\ [@specialized_data_request].

#### Step-by-Step Guide

::: {style="margin-left: 2rem"}
1. If you have not already, complete the process to get access to Research Plus and NCCR Data\ [@specialized_data_request].

2. Select the dataset(s) you would like to gain access to from the [list of available](https://seer.cancer.gov/data/specialized/available-databases/). Be sure to review the documentation associated with the data you want to request\ [@specialized_data; @specialized_data_request].

3. Log into the [SEER Specialized Database](https://seer.cancer.gov/data/specialized/wayf/) to submit your data request\ [@specialized_data_login; @specialized_data_request].
:::

Additional details pertaining to the data application and application review process can be found in the Databases DAC [Standard Operating Procedures (SOP)](https://seer.cancer.gov/data/SEER-Specialized-DBs-DAC-SOP.pdf)\ [@specialized_data_request].

:::

## Valuable Links

::: {style="margin-left: 2rem"}
- [Documentation for SEER Data](https://seer.cancer.gov/data-software/documentation/): Find details about SEER databases, the NCCR databases, and variable definitions available in SEER*Stat, including record numbers, U.S. population coverage, geographic regions, and SEER variable dictionary\ [@data_docs].

- [More About the Data](https://seer.cancer.gov/data/): Find information on SEER Incidence Database, data access, policies, FAQs, data history, specialized databases, and contacts\ [@data_comparisons].

- [About the SEER Registries](https://seer.cancer.gov/registries/): Find additional details about participating registeries, what role they play in the data collection process, and the demographic representation they the sites include\ [@registries].

- [Methods & Tools for Population-based Cancer Statistics](https://surveillance.cancer.gov/tools/): Find documentation about the statistical methods implemented in SEER*Stat software that generate the various outputs listed above in the overview\ [@methods_tools].

- [Supplementary Datasets](https://seer.cancer.gov/data-software/datasets.html): Find additional details about data commonly used with SEER or NCCR, including U.S. Mortality, U.S. Populations, Standard Populations, County/Tract Attributes, Expected Survival, and SEER Linked Databases. The Linked Databases include SEER-Medicare, SEER-Medicare Health Outcomes Survey (SEER-MHOS), SEER-Medicaid, and SEER-Consumer Assessment of Healthcare Providers and Systems (SEER-CAHPS)\ [@us_data].
:::


## Publications

```{r}
#| echo: false

# Fetch and format the citations.
citations <- get_citations(
  doi = c("10.1016/j.urology.2022.07.012", "10.1016/j.annonc.2022.03.276", 
          "10.1016/j.jgo.2022.09.010", "10.1016/j.eclinm.2023.102138", 
          "10.59249/IAJU7580"),
  affiliation = c("Yale School of Public Health", "Yale School of Medicine", "Yale University"),
  highlight_names = c("Bhramar Mukherjee")
)

# Condition the output based on presence of articles.
citations_html <- make_pub_list(citations)

if(!any(str_detect(citations$citations, "No publications"))) {
  citation_length <- length(citations$citations)
  citation_date   <- str_match(citations$query, "\\((\\d{4}):(\\d{4})\\[dp\\]\\)")[,2]
} else {
  citation_length <- 0
  citation_date   <- as.numeric(format(Sys.Date(), "%Y")) - as.numeric(str_extract("No publications found in the last 25 years.", "[0-9]{2}"))
}

```


This section presents a selection of PubMed articles that utilize the dataset and are authored by individuals affiliated with the Yale School of Public Health. These articles are provided to inspire researchers and students to use the data in their own work.

::: {style="margin-left: 1rem"}

```{r}
#| echo: false
#| results: 'asis'

# Display formatted citations.
cat(citations_html)
```

```{=html}
<!-- Pagination buttons and scripting -->
<div class="citations-navigation">
  <button onclick="prevPage()">&#9664; Previous</button>
  <button onclick="nextPage()">Next &#9654;</button>
</div>

<script src="../../../Code/pagination.js"></script>
```

:::