---
title-block-banner: true
title: |
  Worldwide Index of Serotype Specific Pneumococcal Antibody Responses (WISSPAR)
description: |
  </br>
  The Worldwide Index of Serotype Specific Pneumococcal Antibody Responses (WISSPAR) is a searchable database launched in 2023 by Yale researchers to aggregate immunogenicity data from pneumococcal vaccine clinical trials. WISSPAR enables researchers and regulators to quickly access and visualize serotype-specific immune response data through search and filtering tools, supporting comparative efficacy research and evidence-based decision making for pneumococcal conjugate vaccine development\ [@perniciaro2023; @clinical_trials].
include-in-header: ../../../_includes/feedback_popup.html
bibliography: BibTeX/yale_freely_references.bib
csl: ../../../Code/nature.csl
---

```{r}
#| echo: false

# Load in the functions.
source('../../../Code/PubMed Search_Functions.R')
```

<small>Updated: January 15<sup>th</sup>, 2026</small>

## Overview

[Worldwide Index of Serotype Specific Pneumococcal Antibody Responses (WISSPAR)](https://wisspar.com/) was created by [Stephanie Perniciaro, PhD, MPH](https://ysph.yale.edu/profile/stephanie-perniciaro/), Dominic Cooper-Wootton, Maria Deloria Knoll, PhD, and [Daniel Weinberger, PhD](https://ysph.yale.edu/profile/daniel-weinberger/) and launched in 2023 as a searchable database of immunogenicity data from pneumococcal vaccine clinical trials. The platform empowers researchers and regulators to quickly access and compare efficacy data for new pneumococcal conjugate vaccines (PCVs) through search and visualization tools\ [@perniciaro2023]. Users can search or visualize trial data by trial number, vaccine products, manufacturers, dosing schedules, geographic regions, immunocompromised populations, or age groups\ [@clinical_trials].

WISSPAR currently offers data from over 70 clinical trials that generated serotype-specific immunogenicity measures published in the Outcome Measures subsection on clinicaltrials.gov\ [@clinical_trials]. Some datasets also include adverse events reporting, subject demographics, and retention pattern information. Data undergoes manual validation for completeness, clarity, and consistent formatting, though users should be aware that inaccurately reported results on clinicaltrials.gov are not corrected through this process\ [@perniciaro2023].

The resource focuses on two measures of serotype-specific immunogenicity: geometric mean concentrations of immunoglobulin G levels measured by enzyme-linked immunosorbent assay (ELISA) or electrochemiluminescence (ECL) assay, and geometric mean titers measured with opsonophagocytic assays (OPA). These blood serum tests provide follow-up results after pneumococcal vaccine dosing, with follow-up timing that may vary across trials. While these assays are largely cross-trial comparable, researchers should be aware that differences between assay types can cause variations that limit direct comparability\ [@perniciaro2023].

## Gaining Access

All data are freely accessible for direct download or online visualization through the platform. Simply go to the data [Export](https://wisspar.com/export) page to directly download, or create quick, customized visualizations of the data on the [Graphical View](https://wisspar.com/pcv_antibodies) page\ [@data_export; @graphical_view].

## Relevant Links

::: {style="margin-left: 2rem"}
- [Export](https://wisspar.com/export): Access and download data directly\ [@data_export].

- [Graphical View](https://wisspar.com/pcv_antibodies): Generate customized visualizations using data available on the platform\ [@graphical_view].

- [Clinical Trials](https://wisspar.com/trial-dashboard): Find a listing of all currently available trials, and filter or search by vaccine products, manufacturers, dosing schedules, geographic regions, immunocompromised populations, or age groups\ [@clinical_trials].

- [Worldwide Index of Serotype-Specific Pneumococcal Antibody Responses (WISSPAR): A curated database of clinical trial data](https://pmc.ncbi.nlm.nih.gov/articles/PMC10409981/): Find the article introducing the WISSPAR resource, which highlights the scientific need being addressed and other important features\ [@perniciaro2023].
:::


## Publications

```{r}
#| echo: false

# # Fetch and format the citations.
# citations <- get_citations(
#   doi = c("10.1016/j.amjmed.2024.02.032"),
#   affiliation = c("Yale School of Public Health", "Yale School of Medicine", "Yale University"),
#   highlight_names = c("Bhramar Mukherjee")
# )
# 
# # Condition the output based on presence of articles.
# citations_html <- make_pub_list(citations)
# 
# if(!any(str_detect(citations$citations, "No publications"))) {
#   citation_length <- length(citations$citations)
#   citation_date   <- str_match(citations$query, "\\((\\d{4}):(\\d{4})\\[dp\\]\\)")[,2]
# } else {
#   citation_length <- 0
#   citation_date   <- as.numeric(format(Sys.Date(), "%Y")) - as.numeric(str_extract("No publications found in the last 25 years.", "[0-9]{2}"))
# }

```


This section presents a selection of PubMed articles that utilize the dataset and are authored by individuals affiliated with the Yale University. These articles are provided to inspire researchers and students to use the data in their own work.

::: {style="margin-left: 1rem"}

```{r}
#| echo: false
#| results: 'asis'

# # Display formatted citations.
# cat(citations_html)
```

```{=html}
<!-- Pagination buttons and scripting -->
<div class="citations-navigation">
  <button onclick="prevPage()">&#9664; Previous</button>
  <button onclick="nextPage()">Next &#9654;</button>
</div>

<script src="../../../Code/pagination.js"></script>
```

:::

## References
